### 2023-08-26 README.qmd
Initial commit

---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)
```

# Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy

[Suparna Chaudhry](http://www.suparnachaudhry.com/) • Lewis & Clark College  
[Marc Dotson](https://marriottschool.byu.edu/directory/details?id=50683) • Brigham Young University  
[Andrew Heiss](https://www.andrewheiss.com) • Georgia State University

<!-- badges: start -->
<!-- badges: end -->

---

## Abstract

As space for civil society has closed around the world, transnational NGOs have faced a crisis of funding. We explore how NGOs have shifted from traditionally Northern funding sources toward grassroots private philanthropic money. How do individual donors in the US feel about donating to legally besieged NGOs abroad? Do legal restrictions on NGOs influence donors’ decision to donate? We use a conjoint survey experiment to argue that domestic political environments of NGO host countries influence preferences of private donors and that legal crackdowns on NGOs serve as a heuristic of organizational deservingness.

### `r emoji_glue(":zipper_mouth_face::sun_behind_large_cloud::person_walking:")`: Note on "silent skywalk" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [{codename}](http://svmiller.com/codename/) to generate a [NICKA](https://en.wikipedia.org/wiki/List_of_U.S._Department_of_Defense_and_partner_code_names)-style internal-to-us project name that won't change.

```{r generate-code-name, eval=FALSE}
library(codename)
codename_message()
#> code name generated by {codename} v.0.5.0. R version 4.3.0 (2023-04-21).

codename(type = "nicka", seed = "why donors donate") |> 
#> [1] "silent skywalk"
```

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.