### 2023-08-26 index.qmd

---
title: "Why Donors Donate"
author: "Suparna Chaudhry, Marc Dotson, and Andrew Heiss"
format:
  html:
    code-tools: false
    toc: false
---

## Abstract

Abstract goes here

## Code

All the raw code and data for this paper is available in a [GitHub repository](https://github.com/andrewheiss/silent-skywalk).