### 2023-08-28 notebook/results-branding.qmd
Divide results into separate notebook pages

---
title: "H~1~: Branding"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "80%", collapse = TRUE,
  dev = "ragg_png"
)

options(
  digits = 3, width = 120,
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)
```

```{r libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(tidybayes)
library(patchwork)
library(scales)
library(glue)
library(gt)
library(gtExtras)
library(here)

# Helpful functions
label_pp <- label_number(accuracy = 1, scale = 100, 
  suffix = " pp.", style_negative = "minus")

fmt_decimal <- label_number(accuracy = 0.001, style_negative = "minus")

# Targets stuff
tar_config_set(store = here('_targets'),
               script = here('_targets.R'))

tar_load(c(means_features_only))

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))

theme_set(theme_ngo())
```

::: {.callout-tip}
## H~1~: Branding

**Donors will be more likely to donate to Oxfam and Red Cross compared to Amnesty International and Greenpeace.**

*Mechanism: awareness of need and contentiousness of issue area*
:::

### Estimand

$$
\begin{aligned}
\theta =&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Organization} = \text{Oxfam or Red Cross} \right) \bigr] - \\
&\ \textbf{E}\bigl[ Y_i \mid \operatorname{do}\left( \text{Organization} = \text{Amnesty International or Greenpeace} \right) \bigr]
\end{aligned}
$$

```{r h1-calc-estimands}
h1_mms <- means_features_only %>% 
  group_by(feat_org, .draw) %>% 
  summarize(avg = mean(.epred))

h1_amces <- h1_mms %>% 
  group_by(feat_org) %>% 
  compare_levels(variable = avg, by = feat_org, comparison = "control")
```

```{r plot-h1-mm-amce, fig.width=10, fig.height=4, out.width="100%"}
#| column: page-inset-right
p_h1_mms <- h1_mms %>% 
  ggplot(aes(x = avg, y = feat_org, fill = feat_org)) +
  stat_halfeye() +
  scale_x_continuous(labels = label_percent()) +
  scale_y_discrete(labels = label_wrap(10)) +
  scale_fill_manual(values = clrs$prism[c(1, 3, 5, 7)], guide = "none") +
  labs(
    x = "Overall average predicted probability",
    y = NULL,
    fill = NULL,
    title = "Estimated marginal means"
  )

p_h1_amces <- h1_amces %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_org,
    delim = " - ", 
    names = c("feat_org", "reference_level")
  ) %>% 
  add_row(avg = 0, feat_org = "Amnesty International") %>% 
  ggplot(aes(x = avg, y = feat_org, fill = feat_org)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_pp) +
  scale_y_discrete(labels = label_wrap(10)) +
  scale_fill_manual(values = clrs$prism[c(1, 3, 5, 7)], guide = "none") +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    fill = NULL,
    title = "Average marginal component effects (AMCEs)"
  )

p_h1_mms | p_h1_amces
```

```{r table-h1-mm-amce}
#| column: page-inset-right
h1_tbl_mm <- h1_mms %>% 
  group_by(feat_org) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}<br>{build_ci(.lower, .upper)}")) %>% 
  arrange(desc(feat_org)) %>% 
  select(feat_org, nice)

h1_tbl_amces <- h1_amces %>% 
  group_by(feat_org) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>%
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0)) %>% 
  mutate(nice = glue("{fmt_decimal(y)}<br>{build_ci(ymin, ymax)}")) %>% 
  mutate(feat_org = str_replace(feat_org, " - ", "−<br>")) %>% 
  mutate(across(starts_with("p_"), ~fmt_decimal(.))) %>% 
  arrange(desc(feat_org)) %>% 
  select(contrast = feat_org, amce_nice = nice, p_neq_0)

bind_cols(
  h1_tbl_mm, 
  add_row(h1_tbl_amces, contrast = "*(Reference)*")
) %>% 
  gt() %>% 
  sub_missing(columns = everything(), missing_text = "—") %>% 
  fmt_markdown(columns = c(nice, amce_nice, contrast)) %>%
  cols_align(align = "center", columns = everything()) %>% 
  cols_align(align = "left", columns = c(feat_org, contrast)) %>% 
  cols_label(
    feat_org = "Organization", 
    nice = "Posterior EMM",
    contrast = "Contrast",
    amce_nice = "Posterior AMCE",
    p_neq_0 = "*p*~direction~",
    .fn = md
  ) %>% 
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>% 
  tab_footnote(
    footnote = "Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.",
    locations = cells_column_labels(columns = c(nice, amce_nice))
  ) %>% 
  tab_footnote(
    footnote = md("The probability of direction (*p*~direction~) is the probability that the posterior AMCE is strictly positive or negative—it is the proportion of the posterior AMCE that is the sign of the median."),
    locations = cells_column_labels(columns = p_neq_0)
  ) %>% 
  gt_add_divider(columns = nice, style = "dashed", weight = px(1)) %>% 
  opt_footnote_marks(marks = "standard") %>% 
  opt_horizontal_padding(3) %>% 
  opts_theme()
```