### 2023-08-29 notebook/results.qmd
Add overall AMCEs

---
title: "All AMCEs"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "80%", collapse = TRUE,
  dev = "ragg_png"
)

options(
  digits = 3, width = 120,
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)
```

```{r libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(tidybayes)
library(patchwork)
library(ggforce)
library(scales)
library(glue)
library(gt)
library(gtExtras)
library(here)

# Targets stuff
tar_config_set(store = here('_targets'),
               script = here('_targets.R'))

tar_load(c(means_features_only, ref_grid_features_only))

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))

theme_set(theme_ngo())
```

```{r create-var-lookups}
var_levels <- tibble(
  variable = c("feat_org", "feat_issue", "feat_transp", "feat_acc", "feat_funding", "feat_govt")
) %>% 
  mutate(levels = map(variable, ~{
    x <- ref_grid_features_only$all_combos[[.x]]
    if (is.numeric(x)) {
      ""
    } else if (is.factor(x)) {
      levels(x)
    } else {
      sort(unique(x))
    }
  })) %>% 
  unnest(levels)

# Make a little lookup table for nicer feature labels
var_lookup <- tribble(
  ~variable,      ~variable_nice,
  "feat_org",     "Organizations",
  "feat_issue",   "Issue areas",
  "feat_transp",  "Transparency",
  "feat_acc",     "Accountability",
  "feat_funding", "Funding sources",
  "feat_govt",    "Relationship with host government"
) %>% 
  mutate(amce_var = str_replace(variable, "feat_", "amces_")) %>% 
  mutate(variable_nice = fct_inorder(variable_nice))
```

```{r calc-mms}
mms_org <- means_features_only %>% 
  group_by(feat_org, .draw) %>% 
  summarize(avg = mean(.epred))

mms_issue <- means_features_only %>% 
  group_by(feat_issue, .draw) %>% 
  summarize(avg = mean(.epred))

mms_transp <- means_features_only %>% 
  group_by(feat_transp, .draw) %>% 
  summarize(avg = mean(.epred))

mms_acc <- means_features_only %>% 
  group_by(feat_acc, .draw) %>% 
  summarize(avg = mean(.epred))

mms_funding <- means_features_only %>% 
  group_by(feat_funding, .draw) %>% 
  summarize(avg = mean(.epred))

mms_govt <- means_features_only %>% 
  group_by(feat_govt, .draw) %>% 
  summarize(avg = mean(.epred))
```

```{r calc-amces}
amces_org <- mms_org %>% 
  group_by(feat_org) %>% 
  compare_levels(variable = avg, by = feat_org, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_org,
    delim = " - ", 
    names = c("value", "reference_level")
  ) %>% 
  add_row(avg = 0, value = "Amnesty International")

amces_issue <- mms_issue %>% 
  group_by(feat_issue) %>% 
  compare_levels(variable = avg, by = feat_issue, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_issue,
    delim = " - ", 
    names = c("value", "reference_level")
  ) %>% 
  add_row(avg = 0, value = "Emergency response")

amces_transp <- mms_transp %>% 
  group_by(feat_transp) %>% 
  compare_levels(variable = avg, by = feat_transp, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_transp,
    delim = " - ", 
    names = c("value", "reference_level")
  ) %>% 
  add_row(avg = 0, value = "No")

amces_acc <- mms_acc %>% 
  group_by(feat_acc) %>% 
  compare_levels(variable = avg, by = feat_acc, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_acc,
    delim = " - ", 
    names = c("value", "reference_level")
  ) %>% 
  add_row(avg = 0, value = "No")

amces_funding <- mms_funding %>% 
  group_by(feat_funding) %>% 
  compare_levels(variable = avg, by = feat_funding, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_funding,
    delim = " - ", 
    names = c("value", "reference_level")
  ) %>% 
  add_row(avg = 0, value = "Many small donors")

amces_govt <- mms_govt %>% 
  group_by(feat_govt) %>% 
  compare_levels(variable = avg, by = feat_govt, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_govt,
    delim = " - ", 
    names = c("value", "reference_level")
  ) %>% 
  add_row(avg = 0, value = "Friendly")

amces_all <- bind_rows(
  lst(amces_org, amces_issue, amces_transp, amces_acc, amces_funding, amces_govt),
  .id = "amce_var"
) %>%
  left_join(var_lookup, by = join_by(amce_var)) %>%
  mutate(value = factor(value, levels = unique(var_levels$levels)))
```

```{r plot-amces-all, fig.width=7, fig.height=6, out.width="100%"}
ggplot(
  amces_all,
  aes(x = avg, y = value, fill = variable_nice)
) +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  stat_halfeye(normalize = "groups") +  # Make the heights of the distributions equal within each facet
  guides(fill = "none") +
  facet_col(facets = "variable_nice", scales = "free_y", space = "free") +
  scale_x_continuous(labels = label_pp) +
  scale_fill_manual(values = clrs$prism[1:6]) +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    title = "Posterior AMCEs"
  )
```