### 2023-09-07 basic-amces-mms.qmd
Create categorical column

---
title: Basic polisci-style AMCEs and MMs
format:
  html:
    code-fold: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "80%", collapse = TRUE,
  dev = "ragg_png"
)

options(
  digits = 3, width = 220,
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)
```

```{r libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(cregg)

results_long <- readRDS("original_data.rds") %>% 
  mutate(choice_binary = ifelse(choice == alt, 1, 0)) %>%
  mutate(choice_alt = factor(alt * choice_binary)) %>% 
  mutate(
    feat_transp = fct_relabel(feat_transp, ~paste("Transparency: ", .x)),
    feat_acc = fct_relabel(feat_acc, ~paste("Accountability: ", .x)),
  )
```

The `results_long` data frame contains 1 row per alternative, and features are stored as factors:

```{r show-structure}
results_long %>% count(id)
results_long %>% count(id, task)
```

In political science, we just want the AMCEs and marginal means, so this is all we really do:

```{r calculate-amces}
amces <- amce(
  results_long,
  choice_binary ~ feat_org + feat_issue + feat_transp + feat_acc + feat_funding + feat_govt,
  id = ~id
)

amces
```

```{r plot-amce, fig.width=7, fig.height=5}
plot(amces) + guides(color = "none")
```

```{r calculate-mm}
mms <- mm(
  results_long,
  choice_binary ~ feat_org + feat_issue + feat_transp + feat_acc + feat_funding + feat_govt,
  id = ~id
)

mms
```

```{r plot-mm, fig.width=7, fig.height=5}
plot(mms) + guides(color = "none") + geom_vline(xintercept = 0.25, linewidth = 0.25, color = "grey50")
```