### 2023-09-08 notebook/results.qmd
targets-ify predictions and plots

---
title: "All MMs and AMCEs"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "80%", collapse = TRUE,
  dev = "ragg_png"
)

options(
  digits = 3, width = 120,
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)
```

```{r libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(tidybayes)
library(patchwork)
library(ggforce)
library(scales)
library(glue)
library(gt)
library(gtExtras)
library(here)

# Targets stuff
tar_config_set(store = here('_targets'),
               script = here('_targets.R'))

tar_load(c(preds_conditional_treatment_only, grid_treatment_only))
tar_load(c(level_lookup, feature_lookup))

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))

theme_set(theme_ngo())
```

```{r calc-mms}
preds_all <- preds_conditional_treatment_only %>% 
  filter(.category == 0) %>% 
  mutate(.epred = 1 - .epred)

mms_org <- preds_all %>% 
  group_by(feat_org, .draw) %>% 
  summarize(avg = mean(.epred))

mms_issue <- preds_all %>% 
  group_by(feat_issue, .draw) %>% 
  summarize(avg = mean(.epred))

mms_transp <- preds_all %>% 
  group_by(feat_transp, .draw) %>% 
  summarize(avg = mean(.epred))

mms_acc <- preds_all %>% 
  group_by(feat_acc, .draw) %>% 
  summarize(avg = mean(.epred))

mms_funding <- preds_all %>% 
  group_by(feat_funding, .draw) %>% 
  summarize(avg = mean(.epred))

mms_govt <- preds_all %>% 
  group_by(feat_govt, .draw) %>% 
  summarize(avg = mean(.epred))
```

```{r calc-amces}
amces_org <- mms_org %>% 
  group_by(feat_org) %>% 
  compare_levels(variable = avg, by = feat_org, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_org,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_issue <- mms_issue %>% 
  group_by(feat_issue) %>% 
  compare_levels(variable = avg, by = feat_issue, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_issue,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_transp <- mms_transp %>% 
  group_by(feat_transp) %>% 
  compare_levels(variable = avg, by = feat_transp, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_transp,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_acc <- mms_acc %>% 
  group_by(feat_acc) %>% 
  compare_levels(variable = avg, by = feat_acc, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_acc,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_funding <- mms_funding %>% 
  group_by(feat_funding) %>% 
  compare_levels(variable = avg, by = feat_funding, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_funding,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_govt <- mms_govt %>% 
  group_by(feat_govt) %>% 
  compare_levels(variable = avg, by = feat_govt, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feat_govt,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_all <- bind_rows(
  lst(amces_org, amces_issue, amces_transp, amces_acc, amces_funding, amces_govt),
  .id = "amce_var"
) %>%
  left_join(select(feature_lookup, amce_var, feature_nice), by = join_by(amce_var)) %>%
  left_join(
    select(level_lookup, contains("_level")), 
    by = join_by(feature_level)
  )
```

```{r plot-mms-all, fig.width=7, fig.height=6, out.width="100%"}
mms_all <- bind_rows(
  lst(
    mms_org = rename(mms_org, feature_level = feat_org), 
    mms_issue = rename(mms_issue, feature_level = feat_issue), 
    mms_transp = rename(mms_transp, feature_level = feat_transp), 
    mms_acc = rename(mms_acc, feature_level = feat_acc), 
    mms_funding = rename(mms_funding, feature_level = feat_funding), 
    mms_govt = rename(mms_govt, feature_level = feat_govt)
  ),
  .id = "mm_var"
) %>%
  left_join(select(feature_lookup, mm_var, feature_nice), by = join_by(mm_var)) %>%
  left_join(
    select(level_lookup, contains("_level")), 
    by = join_by(feature_level)
  )

ggplot(
  mms_all,
  aes(x = avg, y = feature_short_level, fill = feature_nice)
) +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  stat_halfeye(normalize = "groups") +
  guides(fill = "none") +
  facet_col(facets = "feature_nice", scales = "free_y", space = "free") +
  scale_x_continuous(labels = label_percent()) +
  scale_fill_manual(values = clrs$prism[1:6]) +
  labs(
    x = "Marginal means of probabilities",
    y = NULL,
    title = "Posterior marginal means"
  )
```

```{r plot-amces-all, fig.width=7, fig.height=6, out.width="100%"}
ggplot(
  amces_all,
  aes(x = avg, y = feature_short_level, fill = feature_nice)
) +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  stat_halfeye(normalize = "groups") +  # Make the heights of the distributions equal within each facet
  guides(fill = "none") +
  facet_col(facets = "feature_nice", scales = "free_y", space = "free") +
  scale_x_continuous(labels = label_pp) +
  scale_fill_manual(values = clrs$prism[1:6]) +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    title = "Posterior AMCEs"
  )
```