### 2023-09-08 notebook/survey-irb.qmd

---
title: "Human subjects and IRB approval"
---

This project has received IRB approval from the human subjects research committees at BYU, Christopher Newport, and Georgia State. We have included a generic version of our protocol below.

<p class="text-center">
<a href="files/byu-irb-approval.pdf" class="btn btn-success btn-md" target="_blank" role="button">
{{< iconify fa6-solid circle-check >}} &nbsp; BYU IRB approval
</a> 
&nbsp;
<a href="files/cnu-irb-approval.pdf" class="btn btn-md btn-success" target="_blank" role="button">
{{< iconify fa6-solid circle-check >}} &nbsp; Christopher Newport IRB approval
</a>
&nbsp;
<a href="files/gsu-irb-approval.pdf" class="btn btn-success btn-md" target="_blank" role="button">
{{< iconify fa6-solid circle-check >}} &nbsp; Georgia State IRB approval
</a>
</p>

---

{{< include survey/irb.md >}}