### 2023-09-08 notebook/manuscript.qmd
Add manuscript link

---
title: "Manuscript"
---

Manuscript iframe stuff goes here.