### 2023-09-09 notebook/style-guide.qmd
Add initial style guide

---
title: "Visual style guide"
resources: 
  - "html/linux-libertine-woff/*"
format: 
  html: 
    toc-depth: 4
---

```{r load-libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(scales)

tar_config_set(store = here::here('_targets'),
               script = here::here('_targets.R'))

# Load graphics functions from targets
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))

show_nice_pal <- function(pal, nrow = 1, show_text = TRUE) {
  out <- tibble(x = 1:length(pal),
                color = pal) %>% 
    mutate(x_nice = glue::glue("[{x}]")) %>% 
    mutate(across(c(color, x_nice), ~fct_inorder(.))) %>% 
    ggplot(aes(x = color, y = 1, fill = color)) +
    geom_tile() +
    scale_x_discrete(expand = c(0, 0)) +
    scale_fill_identity() +
    facet_wrap(vars(x_nice), scales = "free_x", nrow = nrow) +
    theme_void(base_family = "Inconsolata") +
    theme(panel.spacing.x = unit(-1, units = "pt"),
          panel.spacing.y = unit(-0.5, units = "lines"),
          strip.text = element_blank())

  if (show_text) {
    out <- out + 
      theme(axis.text.x = element_text(margin = margin(b = 5)),
            strip.text = element_text(size = rel(1.1), margin = margin(t = 5)),
            panel.spacing.y = unit(0, units = "lines"))
  }

  return(out)
}
```

# Fonts

```{scss echo=FALSE}
@import url('https://fonts.googleapis.com/css2?family=Inter&display=swap');

@font-face {
  font-family: 'Linux Libertine O';
  font-style: normal;
  font-weight: normal;
  src: local('Linux Libertine O Regular'), url('/notebook/html/linux-libertine-woff/LinLibertine_R.woff') format('woff');
}

@font-face {
  font-family: 'Linux Libertine O';
  font-style: italic;
  font-weight: normal;
  src: local('Linux Libertine O Italic'), url('/notebook/html/linux-libertine-woff/LinLibertine_RI.woff') format('woff');
}

@font-face {
  font-family: 'Linux Libertine O';
  font-style: normal;
  font-weight: bold;
  src: local('Linux Libertine O Bold'), url('/notebook/html/linux-libertine-woff/LinLibertine_RB.woff') format('woff');
}

@font-face {
  font-family: 'Linux Libertine O Bold';
  font-style: italic;
  font-weight: bold;
  src: local('Linux Libertine O Bold Italic'), url('LinLibertine_RBI.woff') format('woff');
}

.font-example {
  background: #ffffff;
  padding: 1.5em 1.5em 0.75em 1.5em;
  margin-bottom: 1rem;
  border-left: 10px solid #68abb8;
  border-top: 1px solid #68abb8;
  border-right: 1px solid #68abb8;
  border-bottom: 1px solid #68abb8;
  font-size: 1.1em;
}

.inter {
  font-family: "Inter";
}

.libertine {
  font-family: "Linux Libertine O Regular";
  font-size: 1.2em;
}
```

We use [Inter](https://rsms.me/inter/) (accessible for free from [Google Fonts](https://fonts.google.com/specimen/Inter)) as the main font in all our plots.

::: {.font-example .inter}
**Inter**

Whereas recognition of the *inherent dignity* and of the ***equal and inalienable*** rights of all members of the human family is the foundation of freedom, justice and peace in the world…
:::

In the fancy preprint version of our manuscript we use [Linux Libertine](https://en.wikipedia.org/wiki/Linux_Libertine) for the main body font, since it fits with [Libertinus Math](https://github.com/alerque/libertinus) for fancier math typesetting.

::: {.font-example .libertine}
**Linux Libertine**

Whereas recognition of the *inherent dignity* and of the ***equal and inalienable*** rights of all members of the human family is the foundation of freedom, justice and peace in the world…
:::

# Colors

## Palettes

### Prism (`clrs$prism`)

```{r show-prism-short, fig.width=8, fig.height=0.2}
#| out-width: 100%
#| echo: false
clrs$prism %>% show_nice_pal(show_text = FALSE, nrow = 1)
```

Our main palette is the [qualitative Prism palette from CARTOColors](https://carto.com/carto-colors/). CARTOColors palettes were designed for data-focused map visualizations,but can be used in any kind of visualization, similar to [ColorBrewer](http://colorbrewer2.org/).

Access the palette using `clrs$prism`, and access specific colors with indexes like `clrs$prism[8]`:

```{r show-prism, fig.width=8, fig.height=2}
#| out-width: 100%
#| echo: false
clrs$prism %>% show_nice_pal(show_text = TRUE, nrow = 2)
```

```{r hex-prism}
#| code-fold: show
clrs$prism %>% set_names(1:length(.))
```

\

### Single-ish-color sequential palettes

In some cases, like with plots that use `geom_lineribbon()`, we use a range of sequential colors. For these, we use several different [sequential palettes from CARTOColors](https://carto.com/carto-colors/) since they tend to match the color scheme from `clrs$prism`.

### Peach (`clrs$peach`)

```{r show-peach, fig.width=8, fig.height=1}
#| out-width: 100%
#| echo: false
clrs$peach %>% show_nice_pal(show_text = TRUE)
```

```{r hex-peach}
#| code-fold: show
clrs$peach %>% set_names(1:length(.))
```

## Color mappings

TODO: Color sets for specific features/levels/contrasts