### 2023-09-12 notebook/model-definitions.qmd

---
title: "Model definitions"
---

TODO: Formal math for the models

# Model

TODO: Intercept only because (1) LOO is best, and (2) it makes most sense causally since all confounding arrows are deleted due to randomization