### 2023-11-18 manuscript/manuscript.qmd
Add EMM explanation and preregistration

---
title: "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"
short-title: "Why Donors Donate"
date: "2023-11-14"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  Non-governmental organizations (NGOs) are a core component of global governance. Over the last two decades, however, an increasing number of host governments have cracked down on NGOs using a series of laws that adversely affect the availability and use of official aid funds for these organizations. In response to this dramatic shift in the NGO funding landscape, international NGOs (INGOs) have increasingly turned to individual donors to offset the loss of traditional funding sources. Prior research on NGO fundraising has primarily examined how financial transparency and organizational accountability influence individual donor behavior. How do these elements of private donor decision-making interact with structural-level factors such as host country politics and worsening civic environments when deciding to donate? Using a conjoint experiment with likely potential donors in the US, we find that financial transparency and accountability can protect against the dampening effect of host government crackdown and criticism of NGOs. Our results have important implications for how NGOs can adapt to worsening civic environments. 
keywords:
  - philanthropy
  - civil society
  - conjoint experiments
  - donor heuristics
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/british-journal-of-political-science.csl
link-citations: true
word-count: 9324

crossref:
  custom:
    - kind: float
      prefix: Table A
      space-before-numbering: false
      name: Table A
      ref-type: apptbl
      latex-env: apptbl
      latex-list-of-file-extension: lost
      latex-list-of-description: Appendix Table
apptbl-cap-location: top
---

```{r setup, include=FALSE}
if (is.null(knitr::pandoc_to())) {
  fmt_out <- "interactive"
} else {
  fmt_out <- knitr::pandoc_to()
}

knitr::opts_chunk$set(
  echo = FALSE, include = FALSE,
  warning = FALSE, message = FALSE
)

knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "100%", collapse = TRUE
)

options(
  digits = 3, width = 120,
  dplyr.summarise.inform = FALSE
)
```

```{r libraries-data}
library(tidyverse)
library(targets)
library(tidybayes)
library(patchwork)
library(ggforce)
library(ggh4x)
library(scales)
library(glue)
library(gt)
library(gtExtras)
library(modelsummary)
library(here)

# Targets stuff
tar_config_set(
  store = here("_targets"),
  script = here("_targets.R")
)

tar_load(c(
  grid_treatment_only, level_lookup, feature_lookup,
  participant_summary, model_summary_table
))

preds_all <- tar_read(preds_conditional_treatment_only)

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))

theme_set(theme_ngo())

gt_linebreak <- function(x) {
  if (fmt_out == "latex") {
    x %>% 
      str_replace_all("\\[", "{[}") %>% 
      str_replace_all("\\]", "{]}") %>% 
      kableExtra::linebreak(align = "c")
  } else {
    str_replace(x, "\\n", "<br>")
  }
}

fmt_markdown_latex <- function(x, columns) {
  if (fmt_out == "latex") {
    fmt_passthrough(x, columns = columns, escape = FALSE)
  } else {
    fmt_markdown(x, columns = columns)
  }
}

gt_pct <- function(x, page_width = 6) {
  if (knitr::pandoc_to("latex")) {
    paste0(page_width * (x / 100), "in")
  } else {
    gt::pct(x)
  }
}
```

```{r calc-mms}
mms_org <- preds_all %>% 
  group_by(feature_level = feat_org, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_issue <- preds_all %>% 
  group_by(feature_level = feat_issue, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_transp <- preds_all %>% 
  group_by(feature_level = feat_transp, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_acc <- preds_all %>% 
  group_by(feature_level = feat_acc, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_funding <- preds_all %>% 
  group_by(feature_level = feat_funding, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_govt <- preds_all %>% 
  group_by(feature_level = feat_govt, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_all <- bind_rows(
  lst(mms_org, mms_issue, mms_transp, mms_acc, mms_funding, mms_govt),
  .id = "mm_var"
) %>%
  left_join(select(feature_lookup, mm_var, feature_nice_paper), by = join_by(mm_var))

mms_all_medians <- mms_all %>% 
  group_by(feature_nice_paper, feature_level) %>% 
  summarize(
    median_qi(avg),
    p_gt_25 = sum(avg > 0.25) / n()
  ) %>% 
  mutate(p_neq_25 = ifelse(y >= 0.25, p_gt_25, 1 - p_gt_25))

# Make a bunch of differently-formatted numbers
mms <- mms_all_medians %>% 
  mutate(level = janitor::make_clean_names(feature_level)) %>% 
  mutate(
    pd = fmt_decimal2(p_neq_25),
    pct = label_percent(accuracy = 0.1)(y),
    nice_full = glue("{fmt_decimal(y)}; {build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_simple = glue("{fmt_decimal(y)}; {build_ci(ymin, ymax)}"),
    nice_ci = glue("{build_ci(ymin, ymax)}"),
    nice_ci_pd = glue("{build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_ci_pd_prefix = glue("95% cred. int. = {build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_full_prefix = glue("{fmt_decimal(y)}; 95% cred. int. = {build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_br = glue("{fmt_decimal(y)}<br>{build_ci(ymin, ymax)}")
  ) %>% 
  split(~level)

# Marginal means for transparency and accountability together
mms_transp_acc <- preds_all %>% 
  group_by(feat_transp, feat_acc, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_ta <- mms_transp_acc %>% 
  group_by(feat_transp, feat_acc) %>% 
  median_qi(avg) %>% 
  mutate(combined = janitor::make_clean_names(paste0(feat_transp, feat_acc))) %>% 
  mutate(
    pct = label_percent(accuracy = 0.1)(avg),
    nice_ci = glue("{build_ci(.lower, .upper)}")
  ) %>% 
  split(~combined)

# Marginal means for transparency/accountability and crackdown together
mms_transp_crackdown <- preds_all %>% 
  group_by(feat_transp_short, feat_govt, .draw) %>% 
  summarize(avg = mean(.epred))

mms_acc_crackdown <- preds_all %>% 
  group_by(feat_acc_short, feat_govt, .draw) %>% 
  summarize(avg = mean(.epred))

mms_tc <- mms_transp_crackdown %>% 
  group_by(feat_transp_short, feat_govt) %>% 
  median_qi(avg) %>% 
  mutate(combined = janitor::make_clean_names(paste0(feat_transp_short, feat_govt))) %>% 
  mutate(
    pct = label_percent(accuracy = 0.1)(avg),
    nice_ci = glue("{build_ci(.lower, .upper)}")
  ) %>% 
  split(~combined)
```

```{r calc-amces}
amces_org <- mms_org %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_issue <- mms_issue %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_transp <- mms_transp %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_acc <- mms_acc %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_funding <- mms_funding %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_govt <- mms_govt %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_all <- bind_rows(
  lst(amces_org, amces_issue, amces_transp, amces_acc, amces_funding, amces_govt),
  .id = "amce_var"
) %>%
  left_join(select(feature_lookup, amce_var, feature_nice_paper), by = join_by(amce_var)) %>%
  left_join(
    select(level_lookup, contains("_level")), 
    by = join_by(feature_level)
  )

amces_all_medians <- amces_all %>% 
  group_by(feature_nice_paper, feature_level) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>% 
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0))

# Make a bunch of differently-formatted numbers
amce <- amces_all_medians %>% 
  mutate(level = janitor::make_clean_names(feature_level)) %>% 
  mutate(
    pd = fmt_decimal2(p_neq_0),
    pp = fmt_pp_int(y),
    pp_abs = fmt_pp_int(abs(y)),
    nice_full = glue("{fmt_decimal(y)}; {build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_ci_pd = glue("{build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_ci_pd_prefix = glue("95% cred. int. = {build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_full_prefix = glue("{fmt_decimal(y)}; 95% cred. int. = {build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_br = glue("{fmt_decimal(y)}<br>{build_ci(ymin, ymax)}")
  ) %>% 
  split(~level)

amces_transp_crackdown <- mms_transp_crackdown %>% 
  group_by(feat_govt) %>% 
  compare_levels(variable = avg, by = feat_transp_short, comparison = "control")

amces_transp_crackdown_medians <- amces_transp_crackdown %>% 
  group_by(feat_govt, feat_transp_short) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>% 
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0))

amce_tc <- amces_transp_crackdown_medians %>% 
  mutate(level = janitor::make_clean_names(feat_govt)) %>% 
  mutate(
    pd = fmt_decimal2(p_neq_0),
    pp = fmt_pp_int(y),
    pp_abs = fmt_pp_int(abs(y)),
    nice_full = glue("{fmt_decimal(y)}; {build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_ci_pd = glue("{build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_ci_pd_prefix = glue("95% cred. int. = {build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_full_prefix = glue("{fmt_decimal(y)}; 95% cred. int. = {build_ci(ymin, ymax)}; $p_d$ = {pd}"),
    nice_br = glue("{fmt_decimal(y)}<br>{build_ci(ymin, ymax)}")
  ) %>% 
  split(~level)

amces_acc_crackdown_medians <- mms_acc_crackdown %>% 
  group_by(feat_govt) %>% 
  compare_levels(variable = avg, by = feat_acc_short, comparison = "control") %>% 
  group_by(feat_govt, feat_acc_short) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>% 
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0))
```

::: {.content-visible when-format="html"}
# Introduction {.hide}
:::

Over the last two decades, governments globally have increasingly cracked down on NGOs [@ChristensenWeinstein:2013; @DupuyRonPrakash:2016; @Heiss:2017; @GlasiusSchalkDeLange:2020; @BakkeMitchellSmidt:2020; @Chaudhry:2022]. Repressive governments have cracked down on these organizations using both violence, as well as administrative crackdown, which create barriers to funding and advocacy for a variety of NGOs [@DupuyRonPrakash:2016; @Chaudhry:2022]. Policymakers have referred to this phenomenon as the "closing of civic space" or "shrinking civic space" [@CarothersBrechenmacher:2014; @Rutzen:2015] and data from the CIVICUS Civil Society Monitor shows that civil society groups face serious restrictions in more than 60% of countries---117 in total [@CIVICUSMonitor]. These attacks---most of which have been implemented through anti-NGO laws---have adversely affected the ability of both domestic and international NGOs to operate and raise funds. Official aid funds have decreased substantially in countries that have imposed restrictive anti-NGO legislation [@ChaudhryHeiss:2018; @DupuyPrakash:2018] and official aid donors channel fewer funds to contentious issues and instead choose to fund "tame" causes [@ChaudhryHeiss:2018; @Bush:2015; @Herrold:2020; @RightSpringmanWibbels:2022]. In response to this dramatic shift in the NGO funding landscape, international NGOs (INGOs) have increasingly turned to foreign individual donors to offset the loss of funding from government and foundation sources [@BanksHulmeEdwards:2015].

The amount of private foreign aid is not insignificant---from 2000--2010, private giving by U.S. entities towards international causes doubled in real terms [@DesaiKharas:2018, 505]. In 2016, U.S. private philanthropy towards international NGOs amounted to \$22.06 billion [@GivingUSA:2017]. Most notably, a large portion of this growth was driven by individuals making small-scale donations. In 2017, individuals gave \$286.7 billion or approximately 70% of total giving to international affairs. This further increased to \$427.71 billion in 2018, 85.7% of which was through individuals [@GivingUSA:2019]. This turn to private donors opens INGOs up to new dynamics in seeking out funds. Many organizations that had previously relied on governments and foundations had optimized their fundraising strategies for those donors, but now need to tailor their fundraising strategies to private individuals, who behave differently from governments and foundations.

Individual donors have different preferences and rely on signals of effectiveness that differ from the benchmarks that large donor agencies rely on. While INGOs have control over these organizational-level factors such as financial transparency and accountability that motivate donations, the political environment in the host countries INGOs work in---including the global crackdown on civil society that is causing many organizations to seek out private funding in the first place---is a structural factor that INGOs can rarely influence on their own. How do organizational elements of private donor decision-making interact with broader structural limits on INGO fundraising? How do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? How do organizational-level factors such as financial transparency and accountability compare with more structural-level factors such as host country civic environments when deciding to donate?

Existing research shows that organizational accountability and financial transparency play an important role in motivating individual donors to donate to NGOs [@HarrisNeely:2021; @SaxtonKuoHo:2012; @Vaccaro:2006; @VaccaroMadsen:2006; @Tremblay-BoirePrakash:2017; @Becker:2018]. We extend this literature by evaluating how these organization-level characteristics influence private donor behavior in an era of shrinking civic space. In considering this question, we engage with a growing body of research on the consequences of closing civic space on NGOs and donors [@DupuyPrakash:2018; @RightSpringmanWibbels:2022; @FransenDupuyHinfelaar:2021; @SundstromHenrySperling:2022; @Herrold:2020; @ChaudhryHeiss:2021]. We argue that NGO-level organizational factors and domestic political environments of NGO host countries serve as heuristics and can influence foreign private donors' preferences. We fielded a conjoint experiment to a nationally representative sample of 1,016 adults in the US, which allows us to directly compare the relative impact of host government repression of NGOs and organizational financial transparency and accountability on individual donor preferences. The result of this experiment shows that both financial transparency and accountability increase individuals' willingness to donate to NGOs. Conversely, learning about host government criticism and crackdown against NGOs decreases the likelihood of donation by itself; however, organizational features such as financial transparency and accountability can protect against this dampening effect. These effect sizes are substantial: our key treatment effects have a magnitude of around ten percentage points in the predicted probability of choosing to donate to an NGO.

Our findings make two main contributions. While recent research on private foreign aid has examined funding decisions of donor NGOs and foundations,[^1] we know less about the funding choices of the individuals responsible for much of these private aid flows [@DesaiKharas:2018]. The systematic study of private foreign aid is still in its infancy---there is an absence of expansive and high quality data [@GreenhillPrizzonRogerson:2013], and most research has been restricted to elite, high net-worth donors [@USTrust2014]. Our research makes a significant contribution to research on individual giving and indicates the promise and limits of different framing appeals on individual donor preferences.

[^1]: On aid from world's largest NGOs, see @Koch:2009, on private aid flows from NGOs based in the US, see @ButheMajorMello-e-Souza:2012.

Second, research shows that NGO leaders often have difficulty in attracting new donors and also find it challenging to motivate current donors [@GaudreauCao:2015]. Examining organizational-level factors that make the average donor increase their charitable giving, especially when facing restricted civic space abroad, can help NGOs better frame their funding appeals towards these donors. Our results have important implications for INGOs facing increased antagonism by governments abroad and point to possible methods of adapting to worsening or shrinking civic environments.

Below, we summarize existing research on determinants of individual giving to foreign NGOs, laying out our expectations regarding how financial transparency and accountability shape individuals' philanthropic preferences in an environment of closing civic space. We then describe our conjoint experiment and present the results. We conclude with implications for NGOs working in repressive countries.

# What determines individual donor behavior?

Substantial research on NGOs has looked at the motivation of individual donors, including altruism, reputational benefits, and alignment with personal values [@BilodeauSlivinski:1997; @Wiepking:2010; @BekkersWiepking:2011]. However, this research has overwhelmingly looked at individual donor behavior and motivation for giving to local---rather than international---causes [@BekkersWiepking:2011; @Tremblay-BoirePrakash:2017]. Understanding the latter is important as philanthropy to international causes can be harder to motivate because the number of recipients is larger and further removed from the donor [@CasaleBaumann:2015, 100]. Further, as @MicklewrightSchnepf:2009 point out, motivations for donating to ongoing international causes differ from donating money to natural disasters or emergency international causes. Under ordinary circumstances, "donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad" [@Tremblay-BoirePrakash:2017, 644]. However, we lack adequate data and theory regarding the conditions under which private donors engage in increased philanthropy towards INGOs. This gap also exists because conventionally, public donors or official aid donors have been among the main funders of INGOs.

The third wave of democratization, the collapse of the Soviet Union, and the spread of a "pro-NGO norm" led to an increased role and visibility for NGOs. Globally, a large amount of official aid resources were channeled to NGOs focusing on service provision and development. Public and private donors perceived these organizations as more efficient than states with weak institutions and poor governance; there was a high risk that such states would use aid transfers inefficiently [@Dietrich:2013]. Less bureaucratic and more nimble than state institutions, NGOs were seen as key agents of development and democratization [@Herrold:2020, 142--43].

But as INGOs became increasingly powerful actors in global policy circles, they faced a number of criticisms. Many organizations struggle with the conflicting imperatives of being large organizations that can operate efficiently at scale, while maintaining the grassroots contacts and consensus building that contributed to their early successes [@MitchellSchmitzBruno-vanVijfeijken:2020; @Jalali:2008; @Bano:2012; @Rahman:2006]. Many nonprofit scandals involve some form of financial wrongdoing [@PrakashGugerty:2010b]. Globally, around 5% of revenue is lost annually due to fraud [@ACFE:2021]. Misappropriation of assets and embezzlement schemes are among the most common forms of financial fraud [@LamotheTer-MkrtchyanRuddle:2023]. These criticisms and concerns, in turn, have led donors to demand greater NGO responsiveness, especially in terms of financial transparency and accountability.

Private donors, however, do not necessarily seek out complete information about an organization's deservingness and efficiency, since this is a costly and time-consuming process [@CrosonShang:2011; @Tremblay-BoirePrakash:2017]. As such, these donors respond to a range of heuristics and signals when deciding whether to donate to an NGO. Nonprofits fundraising from individual donors may therefore rely on framing to control perceptions [@ChaudhryHeiss:2021]. Below, we theorize the effect of different frames surrounding a country's legal environment and nonprofit organizational characteristics on individual donor preferences.

## Organizational heuristics: The impact of NGO organizational practices on individual philanthropy

In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, individuals rely on heuristics to simplify their decision-making. One heuristic for potential donors is organizational efficacy or the belief that an individual's donation will make a real difference [@ChaudhryHeiss:2021]. This can manifest itself through multiple mechanisms such as the amount an organization spends on overhead costs, whether an organization makes its finances public, or whether an organization participates in voluntary accountability practices, such as listing organizational details on websites like GuideStar and Charity Navigator.

Financial transparency, or the degree of completeness of information provided by organizations to the public concerning their activities, is one tangible heuristic that donors use to make their decisions. NGOs can engage in financial transparency in a variety of ways such as distribution of audited financial statements and third party intermediaries. Research shows that transparency is associated with greater donations, lower debt, and better governance [@HarrisNeely:2021; @SaxtonKuoHo:2012]. It can also enhance donor trust and confidence in organizations [@Vaccaro:2006; @VaccaroMadsen:2006].

Previous research has found that media exposés about charity mismanagement can generate negative reputational spillovers for the charitable sector [@GibelmanGelman:2004]. Diversions, or the unauthorized use of an organization's assets, including embezzlement and theft can result in a decrease in donations. This effect becomes even stronger with media coverage of diversion [@HarrisPetrovitsYetman:2023]. Organizational financial transparency can reduce actual or perceived information asymmetries between donors and charities, thus potentially increasing donor likelihood of donating to an INGO.

> **H~1a~**: If NGOs are financially transparent, then individual private donors will be more likely to donate to them.

Accountability is another factor that can impact donor willingness to donate. Accountability is necessary to promote public trust in NGOs. While transparency is an organization's willingness to make critical data about the organization public, accountability is the willingness of the organization to explain its action to shareholders. It also encompasses how well an organization's mission is aligned with its resources and how it manages resources to optimize performance [@Sloan:2009].

Explaining actions to stakeholders can be important because many NGOs, even when well-intentioned ones, can devote excessive resources towards marketing and overheads [@Bowman:2006]. Private donors, subsequently, may be more likely to dislike NGOs that spend a large share of their revenues for unproductive purposes [@Rose-Ackerman:1982], or rather, a large share of resources on administration, management and fundraising. Lack of accountability is associated with lesser public trust, reputation, perceived quality and donation behavior [@Becker:2018]. As @Tremblay-BoirePrakash:2017 point out, while individuals would like to have some assurance that their resources will be used appropriately and organizations are not spending too much on overhead, donors do not necessarily have the time to research charities thoroughly.

To judge organizational accountability, donors may use ratings by watchdog organizations like Charity Navigator and GuideStar as heuristics to guide their decisions. Research shows that initiatives to address accountability gaps by these organizations can affect donations to organizations in a significant manner [@Sloan:2009; @GordonKnockNeely:2009]

> **H~1b~**: If NGOs are accountable and hold regular third-party audits, then individual private donors will be more likely to donate to them.

## Structural heuristics: Host country conditions

While NGOs have control over organizational factors such as transparency and accountability and can work to improve these characteristics and public perception of them, NGOs may have little to no control over structural factors that can also influence philanthropy. These structural factors include political contexts in NGO host countries, especially their environments for civil society organizations. A global cascade of anti-NGO laws in recent decades has created barriers to entry, funding, and advocacy for civic organizations [@GlasiusSchalkDeLange:2020; @BakkeMitchellSmidt:2020; @Chaudhry:2022; @Heiss:2017; @DupuyRonPrakash:2016]. As a result, NGOs must spend more time, effort, and resources to ensure their survival---at the expense of pursuing their missions.

Some NGOs have adapted to repressive environments by recreating their organizations and changing their mission to avoid directly confronting the government. Analyzing the impact of anti-NGO laws in Bangladesh and Zambia, scholars have found that NGOs with a broad mission ended up restricting themselves to service work and giving up on their advocacy work. Those that continued to engage in transnational advocacy changed the targets of their advocacy, the issues they focused on, and the language they used [@FransenDupuyHinfelaar:2021]. Similarly, in Russia, many NGOs working with foreign partners and lobbying the central government no longer consider it an effective strategy [@SundstromHenrySperling:2022]. Overall, fewer groups may end up focusing on contentious causes, instead focusing on regime-compatible causes.

The changes in civic environments impacting NGO activity have also influenced official donor responses. Scholars have found that while multilateral donors do not reduce aid following the passage of anti-NGO laws, they are associated with a 32% decline in bilateral aid inflows [@DupuyPrakash:2018]. Further, advocacy-oriented donors---i.e., those that fund democracy and civil society promotion activities---reduced their spending by 74%, but did not cut their spending on development projects such as education, health, water, and sanitation [@RightSpringmanWibbels:2022]. However, previous literature has established that individual private donors may not necessarily use the same metrics as official donors when deciding to donate [@DesaiKharas:2018; @ChaudhryHeiss:2021]. Private foreign aid does not have to deal with the same strategic and political considerations as official aid and therefore, may be in a better position to respond to recipients' needs on the grounds [@EasterlyWilliamson:2011]. These anti-NGO laws may signal to private donors that NGOs undertake crucial work abroad, which is why governments perceive them as threatening and crack down on them. Therefore, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization.

> **H~2~**: If NGOs face legal crackdowns abroad, then individual donors will be more likely to donate to them.

## Examining the joint impact of organizational and structural factors

Identifying the causal link between structural factors and donor behavior is more difficult than measuring the link between organizational factors and donor behavior because (1) NGOs have less direct control over their political environments, and (2) organizational characteristics like financial transparency and accountability can sour the NGO-government relationship---governments may be more likely to target NGOs that lack financial transparency and accountability. Governments do not enact anti-NGO laws in isolation. Countries are more likely to repress NGOs when organizational features are in tension with government preferences and when organizations receive substantial funding from abroad [@DupuyRonPrakash:2015]. Therefore, the same organizational characteristic that can increase the likelihood of government crackdown can also simultaneously serve as a donor heuristic and influence perceptions of NGO deservingness.

Donors therefore may show increased willingness to donate to NGOs facing government crackdown/criticism while being financial transparent and accountable. While concerns about NGO mismanagement and lack of financial transparency may signal concerns about an NGO's operations and its ability to abide by the host country's regulatory environment, the presence of such transparency and accountability may instead convince donors that government targeting of NGOs may be ill-intentioned rather than simply a manifestation of financial oversight. Donors may view anti-NGO laws and inhospitable civic environments as a sign of governments looking to crack down on groups that seek to keep governments accountable, rather than merely auditing and punishing groups that violate routine regulations.

As a result, we are uncertain about the interaction between transparency, accountability, and host country legal environment. We suspect that there is a relationship, but we are uncertain about the exact direction or magnitude. Accordingly, we do not propose any confirmatory hypothesis testing, but instead explore the interaction between our primary treatment variables to inform future research on the interplay between organizational and structural dynamics.

Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the country an NGO operates in. The average donor will not know exactly what "crackdown" entails for any given organization (i.e., does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?). For the sake of this experiment, this ambiguity is not a central concern as the heuristics we explore here are explicitly simplified short- cuts to donor information-gathering. Knowing that an organization is financially transparent (even if donors are not familiar with specific kinds of requirements regarding transparency), or knowing that an NGO faces legal trouble abroad should be enough to trigger the effect of these frames on donor preferences. Moreover, although many NGOs make information about their operations accessible through annual reports, it is time-consuming to track down that information. In each of these hypotheses, we assume that donors will not spend additional effort to research the exact details concerning crackdown, transparency and accountability. Instead, donors rely on signals and heuristics about the organization to shape their preferences and behavior [@SzperPrakash:2011].

# Research design

It is possible to study the effects of both organizational and structural heuristics on donor decision making individually, but disentangling the interaction of the two heuristics from individual donor characteristics adds complex dimensionality and makes more standard experimental work difficult, costly, and statistically fraught. To address this, we use conjoint analysis to simultaneously test individual donor preferences in the face of both organizational and structural heuristics. Conjoint analysis is generally applicable whenever individuals face complex, multi-dimensional tradeoffs that make it difficult (or even impossible) to explicitly elicit preferences. Conjoint designs have been used extensively in political science to study complex preferences for elections, laws, public policy, and public opinion [@HainmuellerHopkinsYamamoto:2014; @Mummolo:2016; @KertzerRenshonYarhi-Milo:2021; @BansakHainmuellerHopkins:2021]. @KnudsenJohannesson:2019 argue that research related to respondent opinions requires conjoint analysis because it accounts for the masking effects of factors as well as the complexity of a respondent's judgment of credibility. Because we also want to capture the complexity of a respondent's judgment of philanthropic preferences, we use a conjoint design to evaluate the effect of organizational and structural heuristics on philanthropic preferences.

A conjoint research design also allows us to calculate treatment effects even if respondents do not see every combination of organization attributes, and correspondingly provides a substantial increase in statistical power. In standard factorial experimental designs, participants would ordinarily need to be shown all possible combinations of experimental attributes. In our experiment, we presented respondents with six possible treatments with randomized attributes for each treatment, yielding 576 possible unique combinations of features. Even with respondents answering twelve iterations of the experiment, not every combination was seen. However, as long as all the possible organizational attributes are well randomized and there are no systematic biases toward specific options (i.e. more respondents select the first option because it is the first) or toward earlier iterations of attribute choices (i.e. respondents are more careful and attentive for the first hypothetical organization than the last), we can pool all observations together for specific attributes of interest while marginalizing across all other attributes [@KertzerRenshonYarhi-Milo:2021]. This allows us to (1) estimate the effect of each experimental treatment even if some unique combinations of attributes were unseen, and (2) use a much smaller sample size than would be required in a more traditional factorial design.

While the ability to select key attributes and marginalize over others provides us with analytic flexibility, it also raises possible issues with multiple comparisons, *p*-hacking, and selective cherry-picking [@BansakHainmuellerHopkins:2021], especially given the fact that we have 576 possible combinations of independent variables to explore. As such, before launching the survey experiment, we preregistered a subset of confirmatory and exploratory hypotheses at the Open Science Framework. These hypotheses deal specifically with our key research questions about the effect of transparency, accountability, and crackdown on the propensity to donate, along with comparison treatment effects for organization brand name, issue area, and funding sources. We also examine the interaction between transparency, accountability, and host-country relationships. Our preregistration protocol is available at `ANONYMIZED_URL` and in the appendix, and our data and reproducible code is available at `ANONYMIZED_URL`.[^prereg]

[^prereg]: For narrative purposes, we do not explore all our preregistered hypotheses in this study—many of them are nested inside each other (e.g., the effect of branding and issue area and transparency and crackdown simultaneously). We instead focus here on the simplest, non-nested hypotheses. H~1a~ and H~1b~ correspond to preregistration question #5; H~2~ corresponds to preregistration question #2; our comparison treatments correspond to preregistration questions #1, #3, and #4. Our exploration of the interaction between transparency, accountability, and crackdown corresponds to part 2 of preregistration question #5.

## Sample

We recruited survey participants through Centiment, a commercial online provider of high quality non-probability opt-in survey panels. Centiment ensures panel quality by actively recruiting representative samples of the US population and provides monetary incentives and rewards to participants. To see how varying NGO characteristics influence the decision to donate, our sample was representative of the population of people who are likely to donate to charity. We asked potential participants a screening question about their philanthropic behavior early in the survey---if a participant responded that they give once every few years or never, they were disqualified from the study. Our reported causal effects are therefore not representative of the entire population and instead are only valid for people interested in charitable giving. We also included an attention check question early in the survey and removed respondents who failed the question. Importantly, these screening questions were presented *prior to* the experimental manipulation to avoid post-treatment bias [@MontgomeryNyhanTorres:2018].

Following screening, we received 1,016 viable responses. @apptbl-sample-details-demographics, @apptbl-sample-details-charity, and @apptbl-sample-details-politics provide a detailed breakdown of the individual characteristics of our sample. In general, respondents were well balanced across genders and tended to be older than the national median age of 36---less than 20% of respondents were under 36 years old. The majority of respondents (45%) had never been married, while nearly 40% were married, with 10% divorced. A quarter of the sample had a high school degree, with the remainder attending some level of college or university. The majority (58%) earned less than the national median income (\$61,372). Since we limited the sample to people who frequently donate to charities, most respondents have a favorable opinion of nonprofits and charities, with more than half reporting a 5 or greater (on a 7-point scale) of trust in charities. Nearly 90% of respondents indicated that it is important to trust charities, rating importance of trust at least a 3 on a 7-point scale. The majority of respondents volunteer at least once a year, donate to charity at least once a year, and give between \$1--\$999 annually. Most rarely attend church, but feel that religion is important. Respondents are also engaged civically (three quarters voted in the last election) and follow national news closely (70% follow the news daily). Respondents are fairly balanced ideologically, with the majority reporting a 3, 4, or 5 on a 7-point scale of liberal--conservative political ideology.

::: {.content-visible when-profile="nice"}
\clearpage
:::

## Experimental design

After collecting baseline information on respondent demographics and attitudes toward charity and voluntarism, we presented respondents with three randomly generated hypothetical international nonprofit organizations with six randomly assigned features, along with an option to select none. In this paper, we hypothesize and empirically test the effect of financial transparency, accountability, and host country relationship on the likelihood of donations. In our experiment, we included a few additional treatments to aid with the interpretation and comparison of effect sizes for our treatments of interest. In total, we varied six different organizational and structural attributes that might have an effect on donor behavior: (1) organization name and branding, (2) organization issue area, (3) financial transparency practices, (4) accountability practices, (5) funding sources, and (6) relationship with host government (see @tbl-design).

::: {#tbl-design}

\footnotesize
\renewcommand*{\arraystretch}{1.75}

```{r}
#| label: table-design
#| include: true
transparency <- c("Doesn't engage in transparency", "Engages in transparency")
accountability <- c("Doesn't engage in accountability", "Engages in accountability")
govt <- c("Friendly relationship with government", "Criticized by government", "Under government crackdown")
orgs <- c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross")
issues <- c("Emergency response", "Environment", "Human rights", "Refugee relief")
funding <- c("Funded primarily by many small private donations", "Funded primarily by a handful of wealthy private donors", "Funded primarily by government grants")

tbl_features_levels <- tribble(
  ~type, ~feature, ~levels,
  "H~1a~", "Financial transparency", inline_listify(transparency),
  "H~1b~", "Accountability", inline_listify(accountability),
  "H~2~", "Relationship with host government", inline_listify(govt),
  "Comparison", "Organization", inline_listify(orgs),
  "Comparison", "Issue area", inline_listify(issues),
  "Comparison", "Funding source", inline_listify(funding)
)

tbl_features_levels %>%
  rowwise() %>% 
  mutate(type = ifelse(
    fmt_out == "latex",
    str_replace_all(type, "~([^~]+)~", "\\\\textsubscript{\\1}"),
    type
  )) %>%
  gt() %>%
  cols_label(feature = "Feature/Attribute", levels = "Levels", type = "Purpose") %>%
  cols_width(feature ~ gt_pct(20), levels ~ gt_pct(50), type ~ gt_pct(15)) %>% 
  fmt_markdown_latex(columns = c("type")) %>%
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>%
  opts_theme() |>
  tab_options(
    quarto.disable_processing = TRUE
  )
```

All experimental features and levels included in the conjoint experiment

:::

At the time of designing the study, Charity Navigator---a large database of U.S. nonprofit financial information---categorized international nonprofit activities into four general causes: (1) development, (2) humanitarian relief, (3) international affairs and human rights advocacy, and (4) environment.[^cn] Accordingly, we used four nonprofits that are stereotypical for each cause: the International Committee of the Red Cross (humanitarian relief), Amnesty International (human rights), Oxfam (development), and Greenpeace (environment). We also varied several other features, including four issue areas (emergency response, environmental advocacy, human rights advocacy, and refugee relief), two organizational practices (financial transparency and third-party accountability audits), three funding sources (many small private donations, a handful of wealthy private donors, and government grants), and three relationships with host governments (friendly, criticized by the government, and under government crackdown).

[^cn]: See <http://web.archive.org/web/20170315000746/https://www.charitynavigator.org/index.cfm?bay=search.categories&categoryid=7>. Under the explicitly "International" category, Charity Navigator listed Development and Relief Services; International Peace, Security, and Affairs; and Humanitarian Relief Supplies. The "Environment" category is listed separately but contains many INGOs such as Greenpeace and the World Wide Fund for Nature (WWF). At the time of writing, Charity Navigator had changed its taxonomy to include 18 broader categories.

After presenting respondents with a set of organizations, we asked them which of the three they would be willing to donate to, along with an option for no selection (see @tbl-example for an example). We then repeated the process eleven more times for a total of twelve randomized iterations of hypothetical combinations of attributes, resulting in 12,192 completed experimental tasks (12 iterations × 1,016 respondents).

::: {#tbl-example}

\footnotesize
\renewcommand*{\arraystretch}{1.75}

```{r}
#| label: table-example
#| include: true
withr::with_seed(12345, {
  tbl_example_task <- tribble(
    ~Attribute, ~`Option 1`, ~`Option 2`, ~`Option 3`, ~None,
    "Organization", sample(orgs, 1), sample(orgs, 1), sample(orgs, 1), "—",
    "Issue area", sample(issues, 1), sample(issues, 1), sample(issues, 1), "—",
    "Transparency", sample(transparency, 1), sample(transparency, 1), sample(transparency, 1), "—",
    "Accountability", sample(accountability, 1), sample(accountability, 1), sample(accountability, 1), "—",
    "Funding sources", sample(funding, 1), sample(funding, 1), sample(funding, 1), "—",
    "Relationship with host government", sample(govt, 1), sample(govt, 1), sample(govt, 1), "—"
  )
})

tbl_example_task %>%
  gt() %>%
  cols_align(align = "center", columns = everything()) %>%
  cols_align(align = "left", columns = Attribute) %>%
  cols_width(everything() ~ gt_pct(17)) %>%
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    location = cells_body(columns = Attribute)
  ) %>%
  opts_theme()
```

Example possible experimental task, with three randomized combinations of organizational attributes

:::

## Modeling and estimands

We analyze the results using a multilevel Bayesian multinomial model. Our experimental data has a natural hierarchical structure, with 3 questions nested inside 12 separate experimental tasks, nested inside each of the 1,016 respondents, which lends itself to multilevel modeling. Relying on standard OLS regression can discard valuable information about (1) the relationships and covariance between the different combinations of feature levels offered to respondents, and (2) individual-specific differences in how respondents react to different feature levels [@JensenMarbleScheve:2021]. Since it was impossible for every respondent to see every possible all 12,000 experimental tasks, multilevel modeling allows us to pool together information about respondents with similar characteristics facing similar sets of choices. Moreover, using random respondent effects provides natural regularization and shrinkage for our estimates---experimental tasks that happened to appear more often due to chance will be accounted for and their frequency will not bias the overall causal effect.

We analyze the effect of the various combinations of organizational features and donor characteristics with the following Bayesian multinomial logit model and priors:

$$
\begin{aligned}
&\ \mathrlap{\textbf{Multinomial probability of selection of choice}_i \textbf{ in respondent}_j} \\
\text{Choice}_{i_j} \sim&\ \mathrlap{\operatorname{Categorical}(\{\mu_{1,i_j}, \mu_{2,i_j}, \mu_{3,i_j}\})} \\[10pt]
&\ \mathrlap{\textbf{Model for probability of each option}} \\
\{\mu_{1,i_j}, \mu_{2,i_j}, \mu_{3,i_j}\} =&\ 
  \mathrlap{\begin{aligned}[t]
  & (\beta_0 + b_{0_j}) + \beta_{1, 2, 3} \text{Organization}_{i_j} + \beta_{4, 5, 6} \text{Issue area}_{i_j} + \\
  &\ \beta_7 \text{Transparency}_{i_j} + \beta_8 \text{Accountability}_{i_j} + \\
  &\ \beta_{9, 10} \text{Funding source}_{i_j} + \beta_{11, 12} \text{Government relationship}_{i_j}
  \end{aligned}}\\[5pt]
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) && \text{Respondent-specific offsets from global probability}\\[10pt]
&\ \textbf{Priors} \\
\beta_{0 \dots 12} \sim&\ \mathcal{N} (0, 3) && \text{Prior for choice-level intercept and coefficients} \\
\sigma_0 \sim&\ \operatorname{Exponential}(1) && \text{Prior for between-respondent variability}
\end{aligned}
$$

We do not include any respondent-level covariates beyond the treatment variables. Because this is an experimental design, any statistical confounding is accounted for during the process of randomization and covariates should have no systematic effect on treatment effects. We do not work with the raw results of the model directly (though a complete table of results is included in @apptbl-coefs-orig). Given the conjoint design, we instead create a complete balanced grid of all 576 combinations of feature levels (4 organizations × 4 issues × 2 transparency × 2 accountability × 3 funding × 3 government relationships) and use the model to calculate predicted probabilities of choice selection for each row of possible treatment values. We then collapse this set of predicted probabilities into estimated marginal means (EMMs) for specific features of interested while marginalizing or averaging over all other predicted variables. To calculate the causal effect of a treatment, we find the average marginal component effect (AMCE), which is equivalent to the difference in estimated marginal means for specific feature levels. For example, when estimating the marginal means of the transparency treatment, we find the average predicted probability across the 288 rows of the reference grid where transparency is true and across the 288 rows where transparency is false---the difference between these two estimated marginal means is the AMCE, or the causal effect of the treatment holding all other values constant, on the probability scale.

# Results

We present the posterior distributions of the marginal means and AMCEs for each of our experimental conditions in @fig-all-results and provide posterior medians, credible intervals, and other model diagnostics in @apptbl-all-results and @apptbl-coefs-orig. Because AMCEs are relative statements (i.e. contrasts between one feature level and a reference level), we try to use logical reference levels: for binary treatments like transparency and accountability, we calculate the difference between false and true levels; for ordered treatments like host country relationship, we calculate the differences between different levels of crackdown compared to no crackdown. To avoid imposing an artificial order on other unordered treatment variables, we report both marginal means and AMCEs [@LeeperHoboltTilley:2020].

```{r}
#| label: fig-all-results
#| fig-width: 12
#| fig-height: 6
#| out-width: "100%"
#| fig-cap: "Estimated marginal means and average marginal component effects for conjoint experiment results"
#| column: page
#| include: true
mms_all <- bind_rows(
  lst(mms_org, mms_issue, mms_transp, mms_acc, mms_funding, mms_govt),
  .id = "mm_var"
) %>%
  left_join(select(feature_lookup, mm_var, feature_nice_paper), by = join_by(mm_var)) %>%
  left_join(
    select(level_lookup, contains("_level")), 
    by = join_by(feature_level)
  )

p1 <- ggplot(
  mms_all,
  aes(x = avg, y = feature_short_level, fill = feature_nice_paper)
) +
  stat_halfeye(normalize = "groups") +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  guides(fill = "none") +
  facet_col(facets = "feature_nice_paper", scales = "free_y", space = "free") +
  scale_x_continuous(labels = label_percent()) +
  scale_fill_manual(values = clrs$prism[1:6]) +
  labs(
    x = "Marginal means of probabilities",
    y = NULL,
    title = "Posterior marginal means"
  )

p2 <- ggplot(
  amces_all,
  aes(x = avg, y = feature_short_level, fill = feature_nice_paper)
) +
  stat_halfeye(normalize = "groups") +  # Make the heights of the distributions equal within each facet
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  guides(fill = "none") +
  facet_col(facets = "feature_nice_paper", scales = "free_y", space = "free") +
  scale_x_continuous(labels = label_pp) +
  scale_fill_manual(values = clrs$prism[1:6]) +
  labs(
    x = "Percentage point change in probability of choice selection",
    y = NULL,
    title = "Posterior AMCEs"
  )

p1 | p2
```

We use median values of the posterior marginal means and AMCEs as point estimates and calculate credible intervals using a 95% equal-tailed quantile interval. However, we are generally less concerned with the exact point estimates of our causal effects and instead focus on the direction and relative magnitude of their posterior distributions. For inference, we calculate the probability of direction ($p_d$), or the probability that the posterior AMCE is strictly positive or negative---it is the proportion of the posterior AMCE that is the sign of the median.

## Effect of comparison treatments

To contextualize the magnitude of the causal effects for our hypotheses of interest, we begin our analysis with a brief overview of the effects of organizational characteristics that nonprofits have less direct control over: their brand name, their issue area, and their funding sources. Brand recognition appears to be a powerful heuristic for donor decision making. Respondents were substantially more likely to prefer an organization when it was identified as the Red Cross, with a median posterior marginal mean of `r mms$red_cross$pct` (95% cred. int. = `r mms$red_cross$nice_ci`)---ten percentage points higher than the 25% probability that would be expected if respondents selected an organization at random. When using Amnesty International as the reference category, the AMCE for the Red Cross is `r amce$red_cross$pp` percentage points (`r amce$red_cross$nice_ci_pd_prefix`). Other organizations see much smaller marginal means and AMCEs. Compared to Amnesty International, Greenpeace causes a small `r amce$greenpeace$pp` percentage point increase (`r amce$greenpeace$nice_ci_pd`) and Oxfam causes a small `r amce$oxfam$pp_abs` percentage point decrease (`r amce$oxfam$nice_ci_pd`) in the probability of selecting an organization. The Red Cross brand name heuristic is the strongest of all the experimental treatments and features, signifying the organization's brand power and goodwill among potential donors.

The issue area an organization works on also serves as a heuristic for donors. As seen in the marginal means in @fig-all-results, organizations focused on human rights and emergency response are more popular than those working on issues related to the environment or refugee relief. When using emergency response (the most popular issue) as the reference category for AMCEs, working with human rights causes a `r amce$human_rights$pp_abs` percentage point decrease (`r amce$human_rights$nice_ci_pd`), while environmental and refugee relief issues see `r amce$environment$pp_abs` (`r amce$environment$nice_ci_pd`) and `r amce$refugee_relief$pp_abs` (`r amce$refugee_relief$nice_ci_pd`) percentage point decreases, respectively. To put these effects in context, these effects are smaller than the the Red Cross effect and a little larger than the Oxfam and Greenpeace effects.

Finally, organization funding source also serves as a reliable donor heuristic. Organizations that are funded by many small donors are substantially more popular than those funded by government grants or a small group of wealthy donors---when compared to many small donors, both government and wealthy individual funding cause `r amce$funded_primarily_by_government_grants$pp_abs` and `r amce$funded_primarily_by_a_handful_of_wealthy_private_donors$pp_abs` percentage point decreases in the probability of selection (`r amce$funded_primarily_by_government_grants$nice_ci_pd` and `r amce$funded_primarily_by_a_handful_of_wealthy_private_donors$nice_ci_pd`). These effects follow existing research on donor efficacy [@ChaudhryHeiss:2021]---when donors know that an organization is funded by others like themselves and that the marginal benefit of their individual donation is important, they are more likely to donate. In contrast, when donors know that an organization's funding does not come from the public and instead comes from a few wealthy donors or large government grants, the marginal benefit they receive from donating decreases and they are less likely to donate. Donors therefore appear to be motivated by some degree of personal efficacy.

## Effect of transparency and accountability

Having explored the effects of general organizational characteristics on the propensity to donate, we can test our first hypotheses and examine the effects of our treatments of interest. As seen in the marginal means in @fig-all-results, both transparency and accountability are surprisingly strong signals of organizational deservingness. Respondents strongly prefer organizations that engage in either transparency or accountability---both treatments have a posterior marginal mean of roughly 30% compared to a baseline equally-at-random probability of 25% (transparency: `r mms$transparency_yes$nice_simple`; accountability: `r mms$accountability_yes$nice_simple`). The AMCEs for each treatment show a roughly 10 percentage point increase in the probability of selection compared to organizations that do not engage in transparency or accountability (transparency: `r amce$transparency_yes$nice_full`; accountability: `r amce$accountability_yes$nice_full`). This effect is nearly the same order of magnitude as the Red Cross effect---an NGO that chooses to be more transparent or that takes steps to demonstrate greater accountability can expect to see an increase in the probability of selection equivalent to the boost of the brand name effect associated with the Red Cross. As predicted, we thus find strong support for both H~1a~ and H~1b~: if NGOs are financially transparent or engage in accountability practices, then individual private donors are roughly 10 percentage points more likely to donate to them. Donors appear to reward NGOs for their efforts to disclose their funding and be more transparent.

There is also some evidence that these two effects are equivalent and perhaps interchangeable. @fig-transp-acc (and @apptbl-mms-trans-acc) shows the posterior marginal means for all four combinations of possible levels of transparency and accountability. When respondents see both as "no" simultaneously, the estimated marginal mean is lower than any of the individual treatments in @fig-all-results, at `r mms_ta$transparency_no_accountability_no$pct` (`r mms_ta$transparency_no_accountability_no$nice_ci`), or 9 percentage points lower than the baseline probability of 25%. When either transparency or accountability is set to "yes", the estimated marginal mean is essentially identical at `r mms$transparency_yes$pct` and `r mms$accountability_yes$pct`, respectively. When both treatments are set to "yes", the estimated marginal mean is `r mms_ta$transparency_yes_accountability_yes$pct` (`r mms_ta$transparency_yes_accountability_yes$nice_ci`), which is roughly the same as the overall estimated marginal mean for the Red Cross. Holding all other treatments constant, the effect of transparency and accountability practices on their own are generally the same, and when combined, the overall favorability of the organization increases substantially. This could indicate that donors do not care about which specific type of organizational practice an NGO engages in and that instead they are looking for some sort of signal that the organization is following best practices in transparency or accountability.

```{r}
#| label: fig-transp-acc
#| fig-width: 6
#| fig-height: 2.5
#| out-width: "80%"
#| fig-cap: "Estimated marginal means for the four combinations of transparency and accountability"
#| include: true
mms_trans_acc_all <- mms_transp_acc %>% 
  filter(
    (feat_transp == "Transparency: Yes" & feat_acc == "Accountability: Yes") | 
      (feat_transp == "Transparency: No" & feat_acc == "Accountability: No")) %>% 
  bind_rows(
    mms_transp %>% 
      filter(feature_level == "Transparency: Yes") %>% 
      rename(feat_transp = feature_level) %>% 
      mutate(feat_acc = "Accountability: No"),
    mms_acc %>% 
      filter(feature_level == "Accountability: Yes") %>% 
      rename(feat_acc = feature_level) %>% 
      mutate(feat_transp = "Transparency: No")
  )

mms_trans_acc_all %>% 
  mutate(
    across(c(feat_transp, feat_acc), 
      list(
        grp = ~str_remove(.x, ": .*"),
        short = ~str_remove(.x, ".*: "))
    )) %>% 
  mutate(combined = paste(feat_transp, feat_acc)) %>% 
  ggplot(aes(x = avg, y = feat_transp, fill = combined)) +
  stat_halfeye(normalize = "panels") +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  facet_nested(
    rows = vars(feat_transp_grp, feat_transp_short),
    cols = vars(feat_acc_grp, feat_acc_short),
    switch = "y", scales = "free_y",
    strip = strip_nested(
      background_x = list(element_rect(fill = "grey90"), element_blank()),
      text_x = list(NULL, element_text(hjust = 0.5)),
      by_layer_x = TRUE,
      background_y = list(element_rect(fill = "grey90"), element_blank()),
      text_y = list(NULL, element_text(hjust = 0.5)),
      by_layer_y = TRUE
    )
  ) +
  scale_x_continuous(labels = label_percent()) +
  scale_fill_manual(values = clrs$prism[c(12, 2, 1, 7)], guide = "none") +
  labs(
    x = "Marginal means of probabilities",
    y = NULL
  ) +
  theme_ngo() +
  theme(
    axis.text.y = element_blank(),
    strip.text.y.left = element_text(hjust = 1)
  )
```

## Effect of crackdown

To test our second hypothesis regarding structural heuristics, we explore the effect of an NGO's host country conditions on the propensity to donate. Contrary to our expectations, respondents appear to prefer donating to NGOs that have friendly relationships with their host governments, with a marginal mean of `r mms$friendly_relationship_with_government$pct` (compared to a baseline of 25%; `r mms$friendly_relationship_with_government$nice_ci`). Respondents are less likely to donate when an organization is criticized by its host government, and far less likely when an organization faces crackdown. Compared to other treatments, an organization facing crackdown has roughly the same negative preference as an organization without transparency or accountability measures, while having a friendly relationship is generally equivalent to engaging in transparency or accountability.

In addition to these overall trends in preferences, we can measure the causal effect of moving from friendly NGO–government relations to a more antagonistic relationship. Using friendly relationship as the reference category, facing criticism by the host government causes a `r amce$criticized_by_government$pp_abs` percentage point reduction in the probability of selection (`r amce$criticized_by_government$nice_ci_pd`). Participants respond more strongly as the relationship becomes more conflictual and restricted---an NGO facing crackdown sees a `r amce$under_government_crackdown$pp_abs` percentage point reduction (`r amce$under_government_crackdown$nice_ci_pd`) in support. We thus find strong evidence against H~2~: if NGOs face crackdown abroad, individual donors are *less* likely to donate to them. For context, the causal effect of facing crackdown is the same magnitude as the Red Cross branding effect, but in the opposite direction---donors seem to penalize NGOs facing criticism to the same extent that they reward the Red Cross. This result is surprising, but clarifies findings in previous research. In a similar vignette-based experiment, @ChaudhryHeiss:2021 find no substantial effect of legal crackdown on the likelihood of donation, and any substantive crackdown-related effects are moderated by other treatments like NGO issue area or funding source. This negative effect might also be related to the notion of donor efficacy. When deciding how to maximize the impact or marginal benefit of their individual donation, donors look for signals that their money will make a difference. Transparency and accountability practices, in addition to knowing that other individuals regularly support the organization, all act as heuristics of efficacy. An NGO facing criticism or crackdown abroad, on the other hand, may signal that potential donations could be used for legal fees, or signal that the NGO had perhaps done something to deserve the legal limitations it faces. Crackdown thus serves as a negative heuristic, signaling that donations might not be used as effectively as donors might hope.

## The interaction between transparency, accountability, and crackdown

Financial transparency and accountability practices both have a positive (and likely interchangeable) effect on NGO favorability, while government criticism and crackdown have a negative effect and discourage potential donors from donating. As seen in @fig-transp-crackdown (and @apptbl-transp-crackdown), these heuristics also interact with each other.[^transp-only] The estimated posterior marginal means for different relationships with host governments move in the same direction regardless of whether an organization engages in transparency practices: organizations with friendly relationships with their host governments are most preferred, while organizations under government crackdown are least preferred. Transparency practices in the presence of crackdown seem to have two general effects. First, engaging in transparency offsets most of the negative effect of facing crackdown. The estimated marginal mean for an organization with friendly host government relationships and no transparency is `r mms_tc$no_friendly_relationship_with_government$pct`, which is equivalent to the estimated marginal mean for an organization under government crackdown that does engage in transparency. On average, donors are indifferent to both situations---again, a marginal mean of 25% represents the probability of selecting an organization at random---but the presence of transparency shifts NGOs facing crackdown from negatively preferred to indifferent, while the absence of transparency shifts NGOs with friendly relationships from positively preferred to indifferent. From a practical perspective, this suggests that NGOs under crackdown can focus on improving specific organizational practices to offset the negative signals that accompany their legal difficulties.

Second, worsening host government relationships weaken the positive effect of transparency. On its own, as seen in @fig-all-results, engaging in transparency causes a `r amce$transparency_yes$pp_abs` percentage point increase in the probability of a respondent selecting an organization. The right panel of @fig-transp-crackdown (and @apptbl-transp-crackdown) shows how the causal effect of transparency changes across different types of government relationships. These AMCEs represent the difference in the estimated posterior marginal means of the two levels of transparency in the left panel. The effect shrinks as relationships become more negative: under friendly conditions, the median posterior AMCE of transparency is `r amce_tc$friendly_relationship_with_government$pp` percentage points (`r amce_tc$friendly_relationship_with_government$nice_ci_pd`); when an NGO is criticized, the transparency effect is `r amce_tc$criticized_by_government$pp` percentage points (`r amce_tc$criticized_by_government$nice_ci_pd`); when an NGO faces crackdown, the effect drops to `r amce_tc$under_government_crackdown$pp` percentage points (`r amce_tc$under_government_crackdown$nice_ci_pd`). The causal effect remains substantially positive regardless of the relationship---even under the worst conditions, engaging in transparency causes a 9 percentage point boost---suggesting that NGOs facing crackdown can still increase their favorability with donors by signalling their commitment to transparency and accountability.

[^transp-only]: We only show the interaction of transparency and crackdown in @fig-transp-crackdown. Given the near interchangeability of the two treatments, the interaction between accountability and crackdown is nearly identical. We include both transparency and accountability in @apptbl-transp-crackdown.

```{r}
#| label: fig-transp-crackdown
#| fig-width: 7.5
#| fig-height: 3
#| out-width: "100%"
#| fig-cap: "Estimated marginal means and average marginal component effects for the interaction between transparency and crackdown"
#| include: true
p1 <- mms_transp_crackdown %>% 
  ggplot(aes(x = avg, y = feat_govt, fill = feat_transp_short)) +
  stat_halfeye() +
  geom_vline(xintercept = 0.25, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) +
  scale_x_continuous(labels = label_percent()) +
  scale_y_discrete(labels = label_wrap(20)) +
  scale_fill_manual(values = c(colorspace::lighten(clrs$prism[1], 0.6), clrs$prism[1])) +
  labs(
    x = "Marginal means of probabilities",
    y = NULL,
    title = "Posterior marginal means",
    fill = "Transparency"
  ) +
  theme_ngo()

p2 <- amces_transp_crackdown %>% 
  ungroup() %>% 
  separate_wider_delim(
    feat_transp_short,
    delim = " - ", 
    names = c("feat_transp_short", "reference_level")
  ) %>% 
  mutate(feat_transp_short = factor(feat_transp_short, levels = levels(mms_transp_crackdown$feat_transp_short))) %>% 
  mutate(feat_govt = factor(feat_govt, levels = levels(mms_transp_crackdown$feat_govt))) %>% 
  ggplot(aes(x = avg, y = feat_govt, fill = "AMCE for transparency")) +
  stat_halfeye() +
  geom_vline(xintercept = 0, color = clrs$prism[8], linetype = "dashed", linewidth = 0.25) + 
  scale_x_continuous(labels = label_pp) +
  scale_y_discrete(labels = label_wrap(20)) +
  scale_fill_manual(values = clrs$prism[7]) +
  labs(
    x = "Percentage point change in\nprobability of choice selection",
    y = NULL,
    title = "Posterior AMCEs",
    fill = NULL
  ) +
  theme_ngo()

p1 | p2
```

# Conclusion

Over recent years, governments globally have systematically controlled and repressed NGOs using both violent tactics and legal means---a phenomenon known as "closing" or "shrinking" civic space. This trend of closing civic space has important implications for local and international NGOs as well as donors. Recent research has analyzed the impact of this crackdown on public or official aid donors [@DupuyPrakash:2018; @Herrold:2020; @RightSpringmanWibbels:2022]. However, we know relatively little about how state repression of NGOs affects the preferences of foreign private donors [@ChaudhryHeiss:2021]. Unlike state donors, individual and elite donors may not necessarily withdraw or reduce support for NGOs facing harassment abroad. This is important because private philanthropy towards organizations working in international affairs continues to grow despite inflation---from 2021 to 2022, individual giving to nonprofits in international affairs grew by over 10.9% in the US [@GivingUSA:2023].

Given the increasing hurdles faced by these nonprofits, how do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? How do organizational-level factors such as financial transparency and accountability compare with more structural-level factors such as host country civic environments when deciding to donate? We argue that informational heuristics about NGOs’ legal and political environments and organizational-level attributes can influence donor preferences for international giving. Using a conjoint experiment with a nationally representative sample of the U.S. population, we find that both financial transparency and accountability increases individuals' willingness to donate NGOs. Learning about host government criticism and crackdown against NGOs decreases the likelihood of donation by itself; however, organizational features such as financial transparency and accountability protect against this dampening effect, increasing the probability of philanthropic domations by nine percentage points under the worst conditions of legal crackdown. Our results highlight the importance of organizational characteristics like transparency and accountability even in an era of closing civic space.

These results have important implications for NGO operations, fundraising, and survivability at a time when many INGOs are dealing with hostile host governments. A comparison of the effects of these organizational and structural factors is critical to NGO fundraising---while NGOs have control over transparency and accountability, as well as public perception of these organizational characteristics, they may little to no control over host government crackdowns over their organizations. As a result, they may need to rely entirely on improving individual donor perception of organizational transparency and accountability, its impact on NGO operations and the need for private donor funds at a time of shrinking civic space. This article thus provides insight into the importance of different framing or informational heuristics that can motivate such individual donor giving.

The responses of elite donors and foundations to government crackdown on civil society organizations remains a ripe area for research. Future research should test how well these results map on to elite individual donors, not just the average private donor, as well as determine the extent to which these preferences translate into behavior. Future research should also test the generalizability of these findings to non-American populations. Donors in European countries may have different preferences when choosing to donate to NGOs, as nonprofits in the European sector typically do not rely as much on government funding, and already receive more support from private donors compared to the American context. Government crackdown on civil society organizations is even more challenging for local NGOs in Global South countries, where much government and public support is channeled via service delivery organizations. Local groups tackling contentious issues may not be able to raise funds domestically, as individuals may prefer donating to service organizations focusing on health, education, and poverty [@AbsarCrowRon:2017]. Even local philanthropists may be deterred from giving to organizations focusing on contentious causes such as advocacy, media freedom, and anti-corruption initiatives due to poor tax incentives or a fear of retribution [@Brechenmacher:2017; @DupuyRonPrakash:2015; @Baoumi:2016]. Future research should examine how local NGOs can overcome individual foreign donors’ concerns about host governments' criticisms of and crackdown on these groups.

# References

::: {#refs .bibliography}
:::

\newpage

:::: {#appendix-count}

# Appendix

## Calculating marginal means and AMCEs from a multinomial model using a reference grid

To find the causal effects defined in each of our estimands, we calculate estimated marginal means (EMMs) by finding the fitted probability-scale values for each cell in a balanced reference grid of all 576 possible combinations of feature levels (2 transparency × 2 accountability × 3 government relationships × 4 organizations × 4 issues × 3 funding = 576 rows). We then calculate group averages and contrasts in group averages for each of the features of interest, marginalizing over all other features. At its core, a "marginal mean" refers to the literal mean in the margins in a contingency table of model predictions, and differences in marginal means are equivalent to marginal effects or regression coefficients.

When working with the results for our multinomial regression model, we rely on estimated marginal means rather than raw regression coefficients because of the complexity of the model. As seen in @apptbl-coefs-orig, the model returns three sets of coefficients per conjoint level. Each coefficient shows the shift in probability that someone will choose an organization from column 1, column 2, or column 3, so we get a coefficient for each of those outcomes (or µ1, µ2, and µ3). Under experimental conditions where cells in the contingency table are randomly assigned, it is safe to assume that the cell proportions are equal and then marginalize (i.e. find the average) across the rows or columns. This allows us to take the average of each set of coefficients (e.g. µ1, µ2, and µ3 for "Transparency = Yes") to create a single value per coefficient.

To convert EMMs and AMCEs to a more interpretable probability scale (rather than the original log odds scale), we generate predicted values (marginalized across the three µ terms) for each of the 576 unique combinations of feature levels. @apptbl-grid-example provides an excerpt from this grid, showing six rows where accountability, organization, issue area, and funding are identical and held constant, while transparency and government relations vary.

::: {#apptbl-grid-example}

\footnotesize

```{r}
#| label: table-grid-example
#| include: true
grid_example <- expand_grid(
  feat_transp = c("Transparency: Yes", "Transparency: No"), 
  feat_acc = "Accountability: Yes", 
  feat_govt = c("Friendly relationship with government", "Criticized by government", "Under government crackdown"),
  feat_org = "Red Cross", 
  feat_issue = "Emergency response", 
  feat_funding = "Funded primarily by many small private donations"
)

preds_small <- preds_all %>% 
  group_by(feat_org, feat_issue, feat_transp, feat_acc, feat_funding, feat_govt) %>% 
  summarize(avg = mean(.epred)) %>% 
  inner_join(
    grid_example, 
    by = join_by(feat_org, feat_issue, feat_transp, feat_acc, feat_funding, feat_govt)
  ) %>% 
  mutate(avg_nice = label_comma(accuracy = 0.001)(avg))

p_sm <- preds_small %>% 
  group_by(feat_transp) %>% 
  summarize(
    emm_numeric = mean(avg),
    emm = label_comma(accuracy = 0.001)(mean(avg)),
    list = knitr::combine_words(avg_nice, sep = " + ", and = "")
  ) %>% 
  mutate(feat_transp = janitor::make_clean_names(feat_transp)) %>% 
  split(~feat_transp)

amce_fake <- p_sm$transparency_yes$emm_numeric - p_sm$transparency_no$emm_numeric
amce_tr_fake <- label_comma(accuracy = 0.001)(amce_fake)
amce_tr_fake_pp <- label_comma(accuracy = 1, scale = 100)(amce_fake)

preds_small %>% 
  select(-avg) %>% 
  ungroup() %>% 
  gt() %>% 
  cols_label(
    feat_transp = "Transparency", 
    feat_acc = "Accountability", 
    feat_govt = "Government",
    feat_org = "Organization",
    feat_issue = "Issue",
    feat_funding = "Funding",
    avg_nice = "EMM"
  ) %>%
  cols_width(everything() ~ gt_pct(12)) %>% 
  cols_width(avg_nice ~ gt_pct(7)) %>% 
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>%
  opts_theme() |>
  tab_options(
    quarto.disable_processing = TRUE
  )
```

Excerpt from complete reference grid of all 576 possible combinations of attribute features and levels

:::

To calculate the marginal mean for a feature, we find the average predicted value across each the levels of that feature. To illustrate, assume that @apptbl-grid-example represents the full reference grid of all experimental features and levels. The marginal means for transparency would be (`r p_sm$transparency_no$list`)/3 = `r p_sm$transparency_no$emm` when transparency is set to "no", and (`r p_sm$transparency_yes$list`)/3 = `r p_sm$transparency_yes$emm` when transparency is set to "yes". In reality, the marginal mean for transparency reported in @fig-all-results and @apptbl-all-results represent the average of 288 rows where transparency is no and 288 rows where transparency is yes.

To calculate the AMCE for a feature, we find the difference in estimated marginal means. If we again assume that @apptbl-grid-example contains the full reference grid, the AMCE for transparency would be `r p_sm$transparency_yes$emm` − `r p_sm$transparency_no$emm`, or `r amce_tr_fake`, or `r amce_tr_fake_pp` percentage points. Again, this is not actually the true causal effect—the real AMCE in @fig-all-results and @apptbl-all-results is the difference in marginal means for the 288 rows where transparency is no and the 288 rows where transparency is yes.

## Additional tables

```{=latex}
\floatstyle{plaintop}
\restylefloat{apptbl}
```

::: {#apptbl-sample-details-demographics}

\footnotesize

```{r}
#| label: table-sample-details-demographics
#| include: true
tbl_participant_summary <- participant_summary %>% 
  select(
    category, 
    Question = clean_name,
    Response = level,
    N = count,
    `%` = nice_prop
  ) %>% 
  split(~category)

tbl_participant_summary$Demographics %>% 
  select(-category) %>% 
  group_by(Question) %>% 
  gt() %>% 
  tab_spanner(label = "Demographics", columns = everything()) %>% 
  cols_width(Response ~ gt_pct(30), everything() ~ gt_pct(15)) %>% 
  cols_align(align = "center", columns = everything()) %>% 
  cols_align(align = "left", columns = Response) %>% 
  opts_theme() |>
  tab_options(
    quarto.disable_processing = TRUE
  )
```

Demographics

:::

\newpage

::: {#apptbl-sample-details-charity}

\footnotesize

```{r}
#| label: table-sample-details-charity
#| include: true
tbl_participant_summary$`Attitudes toward charity` %>% 
  select(-category) %>% 
  group_by(Question) %>% 
  gt() %>% 
  tab_spanner(label = "Attitudes toward charity", columns = everything()) %>% 
  cols_width(Response ~ gt_pct(30), everything() ~ gt_pct(15)) %>% 
  cols_align(align = "center", columns = everything()) %>% 
  cols_align(align = "left", columns = Response) %>% 
  opts_theme() |>
  tab_options(
    quarto.disable_processing = TRUE
  )
```

Attitudes toward charity

:::

::: {#apptbl-sample-details-politics}

\footnotesize

```{r}
#| label: table-sample-details-politics
#| include: true
tbl_participant_summary$`Politics, ideology, and religion` %>% 
  select(-category) %>% 
  group_by(Question) %>% 
  gt() %>% 
  tab_spanner(label = "Politics, ideology, and religion", columns = everything()) %>% 
  cols_width(Response ~ gt_pct(30), everything() ~ gt_pct(15)) %>% 
  cols_align(align = "center", columns = everything()) %>% 
  cols_align(align = "left", columns = Response) %>% 
  opts_theme() |>
  tab_options(
    quarto.disable_processing = TRUE
  )
```

Politics, ideology, and religion

:::

\newpage

::: {#apptbl-all-results}

\scriptsize

```{r}
#| label: table-all-results
#| include: true
mms_all_tbl <- mms_all %>% 
  drop_na(feature_level) %>% 
  group_by(feature_nice_paper, feature_level = fct_rev(feature_level)) %>% 
  median_qi(avg) %>% 
  mutate(nice = glue("{fmt_decimal(avg)}\n{build_ci(.lower, .upper)}")) %>% 
  select(feature_nice_paper, feature_level, mm_nice = nice)

amces_all_tbl <- amces_all %>% 
  mutate(feature_level = factor(feature_level, levels = levels(mms_all$feature_level))) %>% 
  mutate(contrast = glue("{feature_level}−{reference_level}")) %>% 
  group_by(feature_nice_paper, feature_level = fct_rev(feature_level), contrast) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>% 
  ungroup() %>% 
  mutate(is_reference = str_detect(contrast, "−NA")) %>% 
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0)) %>% 
  mutate(contrast = ifelse(is_reference, "*(Reference)*", contrast)) %>% 
  mutate(nice = ifelse(is_reference, NA, glue("{fmt_decimal(y)}\n{build_ci(ymin, ymax)}"))) %>% 
  mutate(pd = ifelse(is_reference, NA, fmt_decimal2(p_neq_0))) %>% 
  select(feature_nice_paper, feature_level, contrast, amce_nice = nice, pd)

left_join(
  mms_all_tbl, amces_all_tbl, 
  by = join_by(feature_nice_paper, feature_level)
) %>% 
  mutate(across(c(feature_level, contrast), ~{
    .x %>% 
      str_replace_all("Transparency: ", "") %>% 
      str_replace_all("Accountability: ", "")
  })) %>% 
  # mutate(contrast = str_replace_all(contrast, "[Transparency: |Accountability: ]", "")) %>% 
  mutate(across(c(mm_nice, amce_nice), ~gt_linebreak(.))) %>% 
  group_by(feature_nice_paper) %>% 
  gt() %>% 
  sub_missing(columns = everything(), missing_text = "—") %>% 
  fmt_markdown_latex(columns = c("mm_nice", "amce_nice")) %>% 
  fmt_markdown(columns = c(contrast)) %>%
  cols_align(align = "center", columns = everything()) %>% 
  cols_align(align = "left", columns = c(feature_level, contrast)) %>% 
  cols_label(
    feature_level = "Feature", 
    mm_nice = "Posterior EMM",
    contrast = "Contrast",
    amce_nice = "Posterior AMCE",
    pd = "*p*~direction~",
    .fn = md
  ) %>% 
  cols_width(
    feature_level ~ gt_pct(18),
    mm_nice ~ gt_pct(12),
    contrast ~ gt_pct(35),
    amce_nice ~ gt_pct(12),
    pd ~ gt_pct(6)
  ) %>% 
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>% 
  tab_style(
    style = cell_fill(color = "grey90"),
    locations = cells_row_groups()
  ) %>% 
  tab_footnote(
    footnote = "Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.",
    locations = cells_column_labels(columns = c(mm_nice, amce_nice))
  ) %>% 
  tab_footnote(
    footnote = md("The probability of direction (*p*~direction~) is the probability that the posterior AMCE is strictly positive or negative—it is the proportion of the posterior AMCE that is the sign of the median."),
    locations = cells_column_labels(columns = pd)
  ) %>% 
  gt_add_divider(columns = mm_nice, style = "dashed", weight = px(1)) %>% 
  opt_footnote_marks(marks = "standard") %>% 
  opt_horizontal_padding(3) %>% 
  opts_theme() %>% 
  tab_options(
    quarto.disable_processing = TRUE
  )
```

Complete results

:::

\newpage

::: {#apptbl-mms-trans-acc}

\footnotesize

```{r}
#| label: table-mms-trans-acc
#| include: true
mms_trans_acc_all %>% 
  group_by(feat_transp, feat_acc) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}\n{build_ci(.lower, .upper)}")) %>% 
  mutate(across(c(nice), ~gt_linebreak(.))) %>% 
  select(feat_transp, feat_acc, nice) %>% 
  group_by(feat_transp) %>% 
  gt() %>% 
  fmt_markdown_latex(columns = c("nice")) %>% 
  cols_align(align = "center", columns = c(nice)) %>% 
  cols_align(align = "left", columns = c(feat_transp, feat_acc)) %>% 
  cols_label(
    feat_acc = "Features", 
    nice = "Posterior EMM",
  ) %>% 
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>% 
  tab_style(
    style = cell_fill(color = "grey90"),
    locations = cells_row_groups()
  ) %>% 
  tab_footnote(
    footnote = "Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.",
    locations = cells_column_labels(columns = c(nice))
  ) %>% 
  opt_footnote_marks(marks = "standard") %>% 
  opt_horizontal_padding(3) %>% 
  opts_theme() %>% 
  tab_options(
    quarto.disable_processing = TRUE
  )
```

Marginal means for all combinations of transparency and accountability

:::

::: {#apptbl-transp-crackdown}

\footnotesize

```{r}
#| label: table-transp-crackdown
#| include: true

mms_transp_crackdown_tbl <- mms_transp_crackdown %>% 
  group_by(feat_transp_short, feat_govt) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}\n{build_ci(.lower, .upper)}")) %>% 
  ungroup() %>% 
  arrange(desc(feat_govt), desc(feat_transp_short)) %>% 
  select(feat_transp_short, feat_govt, nice)

amces_transp_crackdown_tbl <- amces_transp_crackdown_medians %>% 
  ungroup() %>% 
  mutate(nice = glue("{fmt_decimal(y)}\n{build_ci(ymin, ymax)}")) %>% 
  mutate(feat_transp_short = str_replace(feat_transp_short, " - ", "−")) %>% 
  mutate(across(starts_with("p_"), ~fmt_decimal(.))) %>% 
  arrange(desc(feat_govt)) %>% 
  select(contrast = feat_transp_short, diff_nice = nice, p_neq_0)

mms_acc_crackdown_tbl <- mms_acc_crackdown %>% 
  group_by(feat_acc_short, feat_govt) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}\n{build_ci(.lower, .upper)}")) %>% 
  ungroup() %>% 
  arrange(desc(feat_govt), desc(feat_acc_short)) %>% 
  select(feat_acc_short, feat_govt, nice)

amces_acc_crackdown_tbl <- amces_acc_crackdown_medians %>% 
  ungroup() %>% 
  mutate(nice = glue("{fmt_decimal(y)}\n{build_ci(ymin, ymax)}")) %>% 
  mutate(feat_acc_short = str_replace(feat_acc_short, " - ", "−")) %>% 
  mutate(across(starts_with("p_"), ~fmt_decimal(.))) %>% 
  arrange(desc(feat_govt)) %>% 
  select(contrast = feat_acc_short, diff_nice = nice, p_neq_0)

bind_rows(
  bind_cols(
    mms_transp_crackdown_tbl, 
    amces_transp_crackdown_tbl %>% 
      add_row(contrast = NA, .after = 1) %>% 
      add_row(contrast = NA, .after = 3) %>% 
      add_row(contrast = NA, .after = 5)
  ) %>% 
    mutate(feat_level = paste("Transparency:", feat_transp_short)) %>% 
    select(-feat_transp_short),
  bind_cols(
    mms_acc_crackdown_tbl, 
    amces_acc_crackdown_tbl %>% 
      add_row(contrast = NA, .after = 1) %>% 
      add_row(contrast = NA, .after = 3) %>% 
      add_row(contrast = NA, .after = 5)
  ) %>% 
    mutate(feat_level = paste("Accountability:", feat_acc_short)) %>% 
    select(-feat_acc_short)
) %>%
  mutate(contrast = ifelse(is.na(contrast), "*(Reference)*", contrast)) %>% 
  mutate(across(c(nice, diff_nice), ~gt_linebreak(.))) %>% 
  relocate(feat_level) %>% 
  mutate(feat_govt = fct_relabel(feat_govt, ~paste("Relationship with government:", .x))) %>% 
  group_by(feat_govt) %>% 
  gt() %>% 
  sub_missing(columns = everything(), missing_text = "—") %>% 
  fmt_markdown_latex(columns = c("nice", "diff_nice")) %>% 
  fmt_markdown(columns = contrast) %>% 
  cols_align(align = "center", columns = everything()) %>% 
  cols_align(align = "left", columns = c(feat_level, contrast)) %>% 
  cols_label(
    feat_level = "Feature", 
    nice = "Posterior EMM",
    contrast = "Contrast",
    diff_nice = "Posterior ∆",
    p_neq_0 = "*p*~direction~",
    .fn = md
  ) %>% 
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>% 
  tab_style(
    style = cell_fill(color = "grey90"),
    locations = cells_row_groups()
  ) %>% 
  tab_footnote(
    footnote = "Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.",
    locations = cells_column_labels(columns = c(nice, diff_nice))
  ) %>% 
  tab_footnote(
    footnote = md("The probability of direction (*p*~direction~) is the probability that the posterior difference in marginal means is strictly positive or negative—it is the proportion of the posterior difference in marginal means that is the sign of the median."),
    locations = cells_column_labels(columns = p_neq_0)
  ) %>% 
  gt_add_divider(columns = nice, style = "dashed", weight = px(1)) %>% 
  opt_footnote_marks(marks = "standard") %>% 
  opt_horizontal_padding(3) %>% 
  opts_theme() %>% 
  tab_options(
    quarto.disable_processing = TRUE
  )
```

Marginal means and AMCEs for interaction between transparency, accountability, and government relationships

:::

\blandscape

::: {#apptbl-coefs-orig}

\footnotesize

```{r}
#| label: table-coefs-orig
#| include: true
notes <- c(
  "Estimates are median posterior log odds from a multinomial logistic regression model with three possible categories, and the columns for µ1, µ2, and µ3 represent estimates for each of the outcomes;",
  "95% credible intervals (equal-tailed quantile intervals) in brackets."
)

base_table <- modelsummary(
  list("Posterior medians" = model_summary_table, "95% credible intervals" = model_summary_table),
  shape = Feature + term ~ group,
  estimate = c("estimate", "[{conf.low}, {conf.high}]"),
  statistic = NULL,
  output = "gt",
  fmt = fmt_significant(2)
)

base_table$`_data` <- base_table$`_data` %>%
  mutate(Feature = ifelse(!duplicated(Feature), as.character(Feature), ""))

# lol there's a weird bug in {gt} where there's an error if the footnote is a
# vector with multiple elements? So we flatten it when knitting to latex
base_table %>% 
  tab_footnote(
    footnote = ifelse(fmt_out == "latex", paste(notes, collapse = " "), notes)
  ) %>%
  opts_theme() |>
  tab_options(
    quarto.disable_processing = TRUE
  )
```

Original model coefficients

:::

\elandscape

\newpage

::: {.content-visible when-profile="ms"}
\singlespacing
:::

## Preregistration

::: {.callout-important icon=false}
### Consensed preregistration

This is an anonymized and condensed version of our full OSF preregistration protocol.
:::

{{< include text/preregistration.md >}}

::::