### 2023-11-20 index.qmd

---
title: "Why Donors Donate"
author: "Suparna Chaudhry, Marc Dotson, and Andrew Heiss"
format:
  html:
    code-tools: false
    toc: false
---

## Abstract

Non-governmental organizations (NGOs) are a core component of global governance. Over the last two decades, however, many host governments have cracked down on NGOs using a series of laws that adversely affect the availability and use of traditional NGO funding. In response to this dramatic shift in the funding landscape, international NGOs have increasingly turned to individual donors to offset the loss funding. Prior research on NGO fundraising has examined how financial transparency and organizational accountability influence individual donor behavior. How do these elements of private donor decision-making interact with structural-level factors such as worsening host country civic environments when deciding to donate? Using a conjoint experiment with likely donors in the US, we find that transparency and accountability can protect against the dampening effect of host government crackdown and criticism of NGOs. Our results have important implications for how NGOs can adapt to worsening civic environments.

## Code

All the raw code and data for this paper is available in a [GitHub repository](https://github.com/andrewheiss/silent-skywalk).