### 2025-01-22 README.qmd
Update readme

---
title: "Navigating Hostility: The Effect of Nonprofit Transparency and Accountability on Donor Preferences in the Face of Shrinking Civic Space"
format: 
  gfm: 
    wrap: preserve
    toc: true
  typst:
    toc: true
    toc-title: " "
    include-in-header: README_stuff/better-layout.typ
    author:
    - name: Suparna Chaudhry
      affiliation: Lewis and Clark College
      orcid: 0000-0003-3506-7590
      url: https://www.suparnachaudhry.com/
      email: schaudhry@lclark.edu
    - name: Marc Dotson
      affiliation: Utah State University
      orcid: 0000-0001-7846-3660
      url: https://occasionaldivergences.com/
      email: marc.dotson@usu.edu
    - name: Andrew Heiss
      affiliation: Georgia State University
      orcid: 0000-0002-3948-3914
      url: https://www.andrewheiss.com
      email: aheiss@gsu.edu
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)

manuscript_target <- targets::tar_read(manuscript)

yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
title <- yaml_metadata$title
abstract <- yaml_metadata$abstract

if (knitr::pandoc_to() == "typst") {
  # typst doesn't support emoji as of 0.11.0, so we include SVG versions from 
  # Noto Emoji (https://github.com/googlefonts/noto-emoji) instead
  emoji_out <- "![](README_stuff/noto-emoji-svgs/emoji_u1f910.svg){height=1em}![](README_stuff/noto-emoji-svgs/emoji_u1f325.svg){height=1em}![](README_stuff/noto-emoji-svgs/emoji_u1f6b6.svg){height=1em}"
} else {
  emoji_out <- emoji_glue(":zipper_mouth_face::sun_behind_large_cloud::person_walking:")
}
```

::: {.content-visible when-format="markdown"}

---

[Suparna Chaudhry](http://www.suparnachaudhry.com/) • Lewis & Clark College  
[Marc Dotson](https://occasionaldivergences.com/about) • Utah State University  
[Andrew Heiss](https://www.andrewheiss.com) • Georgia State University

---

<!-- badges: start -->
[![Preprint](https://img.shields.io/badge/Preprint-10.31235%2Fosf.io%2Fhu6a5-blue)](https://doi.org/10.31235/osf.io/hu6a5) [![OSF DOI](https://img.shields.io/badge/OSF%20project-10.17605%2FOSF.IO%2FR59XZ-blue)](https://doi.org/10.17605/OSF.IO/R59XZ) [![Code DOI](https://img.shields.io/badge/Code-10.5281%2Fzenodo.14220206-blue)](https://doi.org/10.5281/zenodo.14220206)
<!-- badges: end -->

:::

---

## Abstract

`r abstract`

## `r emoji_out`: Note on "silent skywalk" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [{codename}](http://svmiller.com/codename/) to generate a [NICKA](https://en.wikipedia.org/wiki/List_of_U.S._Department_of_Defense_and_partner_code_names)-style internal-to-us project name that won't change.

```{r generate-code-name, eval=FALSE}
library(codename)
codename_message()
#> code name generated by {codename} v.0.5.0. R version 4.3.0 (2023-04-21).

codename(type = "nicka", seed = "why donors donate")
#> [1] "silent skywalk"
```

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.