### 2025-04-15 manuscript/appendix.qmd
Final edits pre-final submission

---
title: 'Appendix for "Navigating Hostility: The Effect of Nonprofit Transparency and Accountability on Donor Preferences in the Face of Shrinking Civic Space"'
short-title: 'Appendix for "Navigating Hostility"'
date: "2025-04-14"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Utah State University
  url: https://occasionaldivergences.com/
  email: marc.dotson@usu.edu
  orcid: 0000-0001-7846-3660
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
  attributes:
    corresponding: true
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/apa.csl
link-citations: true
crossref: 
  tbl-title: "Table&nbsp;A"
  fig-title: "Figure&nbsp;A"
toc: true
format: 
  html: default
  hikmah-pdf:
    include-in-header: 
      - tex/tbl-fig-to-toc.tex
  hikmah-manuscript-pdf+ms:
    include-in-header: 
      - tex/tbl-fig-to-toc.tex
---

```{r setup, include=FALSE}
if (is.null(knitr::pandoc_to())) {
  fmt_out <- "interactive"
} else {
  fmt_out <- knitr::pandoc_to()
}

knitr::opts_chunk$set(
  echo = FALSE, include = FALSE,
  warning = FALSE, message = FALSE
)

knitr::opts_chunk$set(
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "100%", collapse = TRUE
)

options(
  digits = 3, width = 120,
  dplyr.summarise.inform = FALSE
)
```

```{r libraries-data}
library(tidyverse)
library(targets)
library(tidybayes)
library(scales)
library(glue)
library(tinytable)
library(modelsummary)
library(here)

# Targets stuff
tar_config_set(
  store = here("_targets"),
  script = here("_targets.R")
)

tar_load(c(
  participant_summary, model_summary_table, feature_lookup, level_lookup, 
  cps_diffs, data_sans_conjoint
))

details <- list(
  n = nrow(data_sans_conjoint),
  n_nice = label_comma()(nrow(data_sans_conjoint)),
  combos = 2 * 2 * 3 * 4 * 4 * 3
)

preds_all <- tar_read(preds_conditional_treatment_only)

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))

if (fmt_out == "latex") {
  theme_appendix <- function(x) theme_tt(x, "tabular", style = "tabularray")
  options(tinytable_tt_theme = theme_appendix)
}

tt_md <- function(text) {
  if (fmt_out %in% c("html", "interactive")) {
    map_chr(text, \(x) {
      x %>%
        markdown::mark_html(text = ., template = FALSE) %>%
        trimws() %>%
        str_remove_all("<p>") %>%
        str_remove_all("</p>")
    })
  } else if (fmt_out == "latex") {
    map_chr(text, \(x) {
      x %>%
        markdown::mark_latex(text = ., template = FALSE) %>%
        trimws()
    })
  } else {
    map_chr(text, \(x) x)
  }
}

tt_linebreak <- function(text) {
  if (fmt_out == "latex") {
    map_chr(text, \(x) {
      if (is.na(x)) {
        return(x)
      }
      x %>%
        str_replace_all("\\[", "{[}") %>% 
        str_replace_all("\\]", "{]}") %>% 
        str_replace_all("\\n", r"(\\\\)") %>%
        paste0("{", ., "}")
    })
  } else {
    map_chr(text, \(x) {
      if (is.na(x)) {
        return(x)
      }
      str_replace(x, "\\n", "<br>")
    })
  }
}

get_row_group_index <- function(x) {
  if (is.factor(x)) {
    x <- as.character(x)
  }

  x_rle <- rle(x)
  x_i <- cumsum(c(1, utils::head(x_rle$lengths, -1)))
  x_i_adjusted <- as.numeric(x_i) + 0:(length(x_i) - 1)
  return(x_i_adjusted)
}
```

```{r calc-mms}
mms_org <- preds_all %>% 
  group_by(feature_level = feat_org, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_issue <- preds_all %>% 
  group_by(feature_level = feat_issue, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_transp <- preds_all %>% 
  group_by(feature_level = feat_transp, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_acc <- preds_all %>% 
  group_by(feature_level = feat_acc, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_funding <- preds_all %>% 
  group_by(feature_level = feat_funding, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_govt <- preds_all %>% 
  group_by(feature_level = feat_govt, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_all <- bind_rows(
  lst(mms_org, mms_issue, mms_transp, mms_acc, mms_funding, mms_govt),
  .id = "mm_var"
) %>%
  left_join(select(feature_lookup, mm_var, feature_nice_paper), by = join_by(mm_var))

mms_all_medians <- mms_all %>% 
  group_by(feature_nice_paper, feature_level) %>% 
  summarize(
    median_qi(avg),
    p_gt_25 = sum(avg > 0.25) / n()
  ) %>% 
  mutate(p_neq_25 = ifelse(y >= 0.25, p_gt_25, 1 - p_gt_25))
```

```{r calc-amces}
amces_org <- mms_org %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_issue <- mms_issue %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_transp <- mms_transp %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_acc <- mms_acc %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_funding <- mms_funding %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_govt <- mms_govt %>% 
  group_by(feature_level) %>% 
  compare_levels(variable = avg, by = feature_level, comparison = "control") %>% 
  ungroup() %>% 
    separate_wider_delim(
    feature_level,
    delim = " - ", 
    names = c("feature_level", "reference_level")
  ) %>% 
  add_row(avg = 0, feature_level = unique(.$reference_level))

amces_all <- bind_rows(
  lst(amces_org, amces_issue, amces_transp, amces_acc, amces_funding, amces_govt),
  .id = "amce_var"
) %>%
  left_join(select(feature_lookup, amce_var, feature_nice_paper), by = join_by(amce_var)) %>%
  left_join(
    select(level_lookup, contains("_level")), 
    by = join_by(feature_level)
  )

amces_all_medians <- amces_all %>% 
  group_by(feature_nice_paper, feature_level) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>% 
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0))
```

```{=latex}
\clearpage
```

::: {.content-visible when-meta="appendix-single-spacing"}
```{=latex}
\begin{singlespace}
```
:::

# Respondent demographics

```{r}
#| label: tbl-sample-details-demographics
#| tbl-cap: Respondent demographics
#| tbl-pos: H
#| include: true
tbl_participant_summary <- participant_summary$participant_summary %>% 
  mutate(level = case_match(
    level,
    "Age: Less than median" ~ "Less than 2017 national median (36)",
    "Income: Less than median" ~ "Less than 2017 national median ($61,372)",
    .default = level
  )) %>%
  mutate(level = str_remove(level, "^[^:]+: ")) %>%
  select(
    category, 
    Question = clean_name,
    Response = level,
    N = count,
    `%` = nice_prop
  ) %>% 
  split(~category)

summary_notes <- list(
  `*` = list(i = 0, j = 2, text = tt_md(paste0("N = ", details$n_nice, " respondents.")))
)

inner <- glue("
rowsep=0pt, row{<<header_rows>>}={abovesep=3pt, belowsep=0pt},
cells={font=\\footnotesize}
",
  header_rows = paste0(get_row_group_index(tbl_participant_summary$Demographics$Question) + 2, collapse = ","),
  .open = "<<", .close = ">>"
)

tbl_participant_summary$Demographics %>% 
  select(-category, -Question) %>% 
  tt(width = c(0.6, 0.15, 0.15), notes = summary_notes) %>%
  group_tt(j = list("Demographics" = 1:3)) %>%
  group_tt(
    i = tbl_participant_summary$Demographics$Question, 
    indent = 1
  ) %>%
  style_tt(i = -1, align = "c", bold = TRUE) %>%
  style_tt(j = 1, align = "l") %>%
  style_tt(j = 2:3, align = "c") %>%
  style_tt(
    i = get_row_group_index(tbl_participant_summary$Demographics$Question),
    line = "b", line_width = 0.05, align = "l", bold = TRUE
  ) %>%
  style_tt(
    bootstrap_class = "table table-sm",
    tabularray_inner = inner
  ) %>%
  format_tt(escape = TRUE)
```

::: {.content-visible when-meta="appendix-single-spacing"}
```{=latex}
\end{singlespace}
```
:::

\newpage

```{r}
#| label: tbl-sample-details-charity
#| tbl-cap: Respondent attitudes toward charity
#| include: true

inner <- glue("
rowsep=0pt, row{<<header_rows>>}={abovesep=3pt, belowsep=0pt},
cells={font=\\footnotesize}
",
  header_rows = paste0(get_row_group_index(tbl_participant_summary$`Attitudes toward charity`$Question) + 2, collapse = ","),
  .open = "<<", .close = ">>"
)

tbl_participant_summary$`Attitudes toward charity` %>% 
  select(-category, -Question) %>% 
  tt(width = c(0.6, 0.15, 0.15), notes = summary_notes) %>%
  group_tt(j = list("Attitudes toward charity" = 1:3)) %>%
  group_tt(
    i = tbl_participant_summary$`Attitudes toward charity`$Question, 
    indent = 1
  ) %>%
  style_tt(i = -1, align = "c", bold = TRUE) %>%
  style_tt(j = 1, align = "l") %>%
  style_tt(j = 2:3, align = "c") %>%
  style_tt(
    i = get_row_group_index(tbl_participant_summary$`Attitudes toward charity`$Question),
    line = "b", line_width = 0.05, align = "l", bold = TRUE
  ) %>%
  style_tt(
    bootstrap_class = "table table-sm",
    tabularray_inner = inner
  ) %>%
  format_tt(escape = TRUE)
```

```{r}
#| label: tbl-sample-details-politics
#| tbl-cap: Respondent politics, ideology, and religion
#| include: true

inner <- glue("
rowsep=0pt, row{<<header_rows>>}={abovesep=3pt, belowsep=0pt},
cells={font=\\footnotesize}
",
  header_rows = paste0(get_row_group_index(tbl_participant_summary$`Politics, ideology, and religion`$Question) + 2, collapse = ","),
  .open = "<<", .close = ">>"
)

tbl_participant_summary$`Politics, ideology, and religion` %>% 
  select(-category, -Question) %>%
  tt(width = c(0.6, 0.15, 0.15), notes = summary_notes) %>%
  group_tt(j = list("Politics, ideology, and religion" = 1:3)) %>%
  group_tt(
    i = tbl_participant_summary$`Politics, ideology, and religion`$Question,
    indent = 1
  ) %>%
  style_tt(i = -1, align = "c", bold = TRUE) %>%
  style_tt(j = 1, align = "l") %>%
  style_tt(j = 2:3, align = "c") %>%
  style_tt(
    i = get_row_group_index(tbl_participant_summary$`Politics, ideology, and religion`$Question),
    line = "b", line_width = 0.05, align = "l", bold = TRUE
  ) %>%
  style_tt(
    bootstrap_class = "table table-sm",
    tabularray_inner = inner
  ) %>%
  format_tt(escape = TRUE)
```

\clearpage

```{r}
#| label: tbl-sample-cps
#| tbl-cap: Sample characteristics compared to nationally representative Current Population Survey (CPS) estimates
#| include: true

rows_with_dagger <- cps_diffs %>%
  mutate(row = row_number()) %>%
  filter(cps_prop_outside_ci == TRUE) %>%
  pull(row)

notes <- list(
  `*` = list(i = 0, j = 4, text = tt_md("Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.")),
  `†` = list(i = rows_with_dagger, j = 3, text = tt_md("National value is substantially different from the sample; the 95% posterior credible interval for the difference between the sample and national proportions contains 0.")),
  a = list(i = 1:5, j = 1, text = "Annual Social and Economic Supplement (ASEC) of the Current Population Survey (CPS), March 2019"),
  b = list(i = 6:7, j = 1, text = "Monthly CPS, September 2019"),
  c = list(i = 8, j = 1, text = "Monthly CPS, November 2018")
)

cps_diffs %>%
  mutate(across(c(prop_sample, national_value,), fmt_pct)) %>%
  mutate(diff_ci = glue::glue(
    "{fmt_pp_int(y)}\n[{fmt_pp_int(ymin)}, {fmt_pp_int(ymax)}]"
  )) %>%
  select(
    Variable = variable,
    Sample = prop_sample,
    National = national_value,
    `∆` = diff_ci
  ) %>%
  mutate(`∆` = tt_linebreak(`∆`)) %>%
  tt(width = c(0.45, 0.1, 0.1, 0.15), notes = notes) %>%
  format_tt(j = 4, escape = FALSE) %>%
  format_tt(j = 1:3, escape = TRUE) %>%
  style_tt(j = 1, align = "l") %>%
  style_tt(j = 2:4, align = "c") %>%
  style_tt(
    bootstrap_class = "table table-sm",
    tabularray_inner = "rowsep=2pt, cells={font=\\footnotesize}, row{1}={valign=b}"
  )
```

\clearpage

# Model details

We use Stan [@Stan-2-32-2 v2.32.2; @cmdstan-2-36 v2.36] through R [@rproject-4-4-2 v4.4.2] and {brms} [@Burkner:2017 v2.22] to estimate the model. We simulate 4 MCMC chains with 5,000 draws in each chain, 1,000 of which are used for warmup, resulting in 16,000 (4,000 × 4) draws per model parameter. We assess convergence with visual inspection, and all chains converge.

Complete results from the model, along with posterior predictive checks, goodness-of-fit measures, and model diagnostics---as well as our code and data---are available at a companion statistical analysis compendium at `ANONYMIZED_URL`.

# Model coefficients and estimated marginal means

When working with the results for our multinomial regression model, we rely on estimated marginal means (EMMs) rather than raw regression coefficients because of the complexity of the model. At its core, a "marginal mean" refers to the literal mean in the margins in a contingency table of model predictions, and differences in marginal means are equivalent to marginal effects or regression coefficients.

To find the causal effects defined in each of our estimands, we calculate EMMs by finding the fitted probability-scale values for each cell in a balanced reference grid of all `r details$combos` possible combinations of feature levels (2 transparency × 2 accountability × 3 government relationships × 4 organizations × 4 issues × 3 funding = `r details$combos` rows). We then calculate group averages and contrasts in group averages for each of the features of interest, marginalizing over all other features.

## Raw model coefficients

As seen in @tbl-coefs-orig, the model returns three sets of coefficients per conjoint level. Each coefficient shows the shift in probability that someone will choose an organization from that appears as the first, second, and third option in the experimental task, represented by µ1, µ2, and µ3. Under experimental conditions where all the feature levels are randomly assigned, it is safe to assume that the cell proportions are equal and then marginalize (i.e. find the average) across the rows or columns. This allows us to take the average of each set of coefficients (e.g. µ1, µ2, and µ3 for "Transparency = Yes") to create a single value per coefficient.

```{r}
#| label: tbl-coefs-orig
#| tbl-cap: Original multinomial logistic regression model coefficients
#| include: true
notes <- c(
  paste0("N = ", details$n_nice, " respondents. Estimates are median posterior log odds from a multinomial logistic regression model with three possible categories, and the columns for µ1, µ2, and µ3 represent estimates for each of the outcomes;"),
  "95% credible intervals (equal-tailed quantile intervals) in brackets."
)

inner <- "
rowsep=0pt,
row{odd[2-27]}={abovesep=0pt},
row{even[4-28]}={belowsep=6pt},
cells={font=\\footnotesize}, row{1}={valign=b}
"

modelsummary(
  list("Posterior medians" = model_summary_table),
  shape = Feature + term ~ group,
  estimate = "estimate",
  statistic = "[{conf.low}, {conf.high}]",
  output = "tinytable",
  fmt = fmt_significant(2),
  notes = tt_md(paste(notes, collapse = " "))
) %>%
  style_tt(
    bootstrap_class = "table",
    tabularray_inner = inner
  )
```

## Converting coefficients to estimated marginal means

To convert EMMs and AMCEs to a more interpretable probability scale (rather than the original log odds scale), we generate predicted values (marginalized across the three µ terms) for each of the `r details$combos` unique combinations of feature levels. @tbl-grid-example provides an excerpt from this grid, showing six rows where accountability, organization, issue area, and funding are identical and held constant, while transparency and government relations vary.

```{r}
#| label: tbl-grid-example
#| tbl-cap: Excerpt from complete reference grid of all 576 possible combinations of attribute features and levels
#| include: true
grid_example <- expand_grid(
  feat_transp = c("Transparency: Yes", "Transparency: No"), 
  feat_acc = "Accountability: Yes", 
  feat_govt = c("Friendly relationship with government", "Criticized by government", "Under government crackdown"),
  feat_org = "Red Cross", 
  feat_issue = "Emergency response", 
  feat_funding = "Funded primarily by many small private donations"
)

preds_small <- preds_all %>% 
  group_by(feat_org, feat_issue, feat_transp, feat_acc, feat_funding, feat_govt) %>% 
  summarize(avg = mean(.epred)) %>% 
  inner_join(
    grid_example, 
    by = join_by(feat_org, feat_issue, feat_transp, feat_acc, feat_funding, feat_govt)
  ) %>% 
  mutate(avg_nice = label_comma(accuracy = 0.001)(avg))

p_sm <- preds_small %>% 
  group_by(feat_transp) %>% 
  summarize(
    emm_numeric = mean(avg),
    emm = label_comma(accuracy = 0.001)(mean(avg)),
    list = knitr::combine_words(avg_nice, sep = " + ", and = "")
  ) %>% 
  mutate(feat_transp = janitor::make_clean_names(feat_transp)) %>% 
  split(~feat_transp)

amce_fake <- p_sm$transparency_yes$emm_numeric - p_sm$transparency_no$emm_numeric
amce_tr_fake <- label_comma(accuracy = 0.001)(amce_fake)
amce_tr_fake_pp <- label_comma(accuracy = 1, scale = 100)(amce_fake)

preds_small %>% 
  select(
    Organization = feat_org, Issue = feat_issue, Transparency = feat_transp,
    Accountability = feat_acc, Funding = feat_funding, Government = feat_govt,
    EMM = avg_nice) %>%
  tt(width = c(rep(0.15, 6), 0.07)) %>%
  style_tt(
    align = "l",
    bootstrap_class = "table table-sm",
    tabularray_inner = "rowsep=2pt, cells={font=\\footnotesize}, row{1}={valign=b}"
  )
```

To calculate the marginal mean for a feature, we find the average predicted value across each the levels of that feature. To illustrate, assume that @tbl-grid-example represents the full reference grid of all experimental features and levels. The marginal means for transparency would be (`r p_sm$transparency_no$list`)/3 = `r p_sm$transparency_no$emm` when transparency is set to "no", and (`r p_sm$transparency_yes$list`)/3 = `r p_sm$transparency_yes$emm` when transparency is set to "yes". In reality, the marginal mean for transparency reported in thee paper reflects the average of 288 rows where transparency is no and 288 rows where transparency is yes.

To calculate the AMCE for a feature, we find the difference in estimated marginal means. If we again assume that @tbl-grid-example contains the full reference grid, the AMCE for transparency would be `r p_sm$transparency_yes$emm` − `r p_sm$transparency_no$emm`, or `r amce_tr_fake`, or `r amce_tr_fake_pp` percentage points. Again, this is not actually the true causal effect—the real AMCE for transparency reported in the paper is the difference in marginal means for the 288 rows where transparency is no and the 288 rows where transparency is yes.

In the main paper, we include plots of the marginal means and AMCEs for all experimental features. The tables below correspond to each figure in the paper.

```{r}
#| label: tbl-all-results
#| tbl-cap: "Complete marginal means and AMCEs (see Figure 1 in main paper)"
#| include: true
mms_all_tbl <- mms_all %>% 
  drop_na(feature_level) %>% 
  group_by(feature_nice_paper, feature_level = fct_rev(feature_level)) %>% 
  median_qi(avg) %>% 
  mutate(nice = glue("{fmt_decimal(avg)}\n{build_ci(.lower, .upper)}")) %>% 
  select(feature_nice_paper, feature_level, mm_nice = nice)

amces_all_tbl <- amces_all %>% 
  mutate(feature_level = factor(feature_level, levels = levels(mms_all$feature_level))) %>% 
  mutate(contrast = glue("{feature_level}−{reference_level}")) %>% 
  group_by(feature_nice_paper, feature_level = fct_rev(feature_level), contrast) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>% 
  ungroup() %>% 
  mutate(is_reference = str_detect(contrast, "−NA")) %>% 
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0)) %>% 
  mutate(contrast = ifelse(is_reference, "*(Reference)*", contrast)) %>% 
  mutate(nice = ifelse(is_reference, NA, glue("{fmt_decimal(y)}\n{build_ci(ymin, ymax)}"))) %>% 
  mutate(pd = ifelse(is_reference, NA, fmt_decimal2(p_neq_0))) %>% 
  select(feature_nice_paper, feature_level, contrast, amce_nice = nice, pd)

all_tbl <- left_join(
  mms_all_tbl, amces_all_tbl, 
  by = join_by(feature_nice_paper, feature_level)
) %>% 
  mutate(across(c(feature_level, contrast), ~{
    .x %>% 
      str_replace_all("Transparency: ", "") %>% 
      str_replace_all("Accountability: ", "")
  }))

inner <- glue("
rowsep=1.5pt, row{<<header_rows>>}={abovesep=6pt, belowsep=0pt},
cells={font=\\scriptsize}, row{1}={valign=b}
",
  header_rows = paste0(get_row_group_index(all_tbl$feature_nice_paper) + 1, collapse = ","),
  .open = "<<", .close = ">>"
)

notes <- list(
  `*` = list(i = 0, j = c(2, 4), text = tt_md(paste0("N = ", details$combos, " estimated marginal means collapsed into feature-specific means. Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets."))),
  `†` = list(i = 0, j = 5, text = tt_md("The probability of direction (*p*~direction~) is the probability that the posterior AMCE is strictly positive or negative—it is the proportion of the posterior AMCE that is the sign of the median."))
)

all_tbl %>% 
  select(-feature_nice_paper) %>%
  mutate(across(c(mm_nice, amce_nice), ~tt_linebreak(.))) %>%
  setNames(
    c("Feature", "Posterior EMM", "Contrast", "Posterior AMCE", tt_md("*p*~direction~"))
  ) %>%
  tt(width = c(0.18, 0.18, 0.40, 0.18, 0.06), notes = notes) %>%
  group_tt(
    i = as.character(all_tbl$feature_nice_paper), 
    indent = 0
  ) %>%
  format_tt(replace = "—") %>%
  format_tt(j = 3, markdown = TRUE, replace = "—") %>%
  format_tt(j = c(2, 4, 5), escape = FALSE, replace = "—") %>%
  style_tt(j = c(1, 3), align = "l") %>%
  style_tt(j = c(2, 4, 5), align = "c") %>%
  style_tt(
    i = get_row_group_index(all_tbl$feature_nice_paper),
    # line = "b", line_width = 0.05, line_color = "#d3d8dc", align = "l", bold = TRUE,
    line = "b", line_width = 0.05, align = "l", bold = TRUE
  ) %>%
  style_tt(
    i = tail(get_row_group_index(all_tbl$feature_nice_paper), -1),
    # line = "t", line_width = 0.05, line_color = "#d3d8dc",
    line = "t", line_width = 0.05
  ) %>%
  style_tt(
    bootstrap_class = "table",
    tabularray_inner = inner
  )
```

\newpage

```{r}
#| label: tbl-mms-trans-acc
#| tbl-cap: "Marginal means for all combinations of transparency and accountability (see Figure 2 in main paper)"
#| include: true

mms_transp_acc <- preds_all %>% 
  group_by(feat_transp, feat_acc, .draw) %>% 
  summarize(avg = mean(.epred)) %>% 
  ungroup()

mms_trans_acc_all <- mms_transp_acc %>% 
  filter(
    (feat_transp == "Transparency: Yes" & feat_acc == "Accountability: Yes") | 
      (feat_transp == "Transparency: No" & feat_acc == "Accountability: No")) %>% 
  bind_rows(
    mms_transp %>% 
      filter(feature_level == "Transparency: Yes") %>% 
      rename(feat_transp = feature_level) %>% 
      mutate(feat_acc = "Accountability: No"),
    mms_acc %>% 
      filter(feature_level == "Accountability: Yes") %>% 
      rename(feat_acc = feature_level) %>% 
      mutate(feat_transp = "Transparency: No")
  )

mms_trans_acc_all_tbl <- mms_trans_acc_all %>% 
  group_by(feat_transp, feat_acc) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}\n{build_ci(.lower, .upper)}")) %>% 
  mutate(across(c(nice), ~tt_linebreak(.)))

inner <- glue("
rowsep=1pt, row{<<header_rows>>}={abovesep=6pt, belowsep=0pt},
cells={font=\\footnotesize}, row{1}={valign=b}
",
  header_rows = paste0(get_row_group_index(mms_trans_acc_all_tbl$feat_transp) + 1, collapse = ","),
  .open = "<<", .close = ">>"
)

notes <- list(
  `*` = list(i = 0, j = 2, text = tt_md(paste0("N = ", details$combos, " estimated marginal means collapsed into feature-specific means. Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets.")))
)

mms_trans_acc_all_tbl %>%
  select(Features = feat_acc, `Posterior EMM` = nice) %>% 
  tt(width = c(0.25, 0.2), notes = notes) %>%
  group_tt(
    i = as.character(mms_trans_acc_all_tbl$feat_transp), 
    indent = 0
  ) %>%
  style_tt(
    i = get_row_group_index(mms_trans_acc_all_tbl$feat_transp),
    # line = "b", line_width = 0.05, line_color = "#d3d8dc", align = "l", bold = TRUE,
    line = "b", line_width = 0.05, align = "l", bold = TRUE
  ) %>%
  style_tt(j = 1, align = "l") %>%
  style_tt(j = 2, align = "c") %>%
  style_tt(
    bootstrap_class = "table",
    tabularray_inner = inner
  )
```

\newpage

```{r}
#| label: tbl-transp-crackdown
#| tbl-cap: "Marginal means and AMCEs for interaction between transparency, accountability, and government relationships (see Figure 3 in paper)"
#| include: true

mms_transp_crackdown <- preds_all %>% 
  group_by(feat_transp_short, feat_govt, .draw) %>% 
  summarize(avg = mean(.epred))

mms_transp_crackdown_tbl <- mms_transp_crackdown %>% 
  group_by(feat_transp_short, feat_govt) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}\n{build_ci(.lower, .upper)}")) %>% 
  ungroup() %>% 
  arrange(desc(feat_govt), desc(feat_transp_short)) %>% 
  select(feat_transp_short, feat_govt, nice)

amces_transp_crackdown <- mms_transp_crackdown %>% 
  group_by(feat_govt) %>% 
  compare_levels(variable = avg, by = feat_transp_short, comparison = "control")

amces_transp_crackdown_medians <- amces_transp_crackdown %>% 
  group_by(feat_govt, feat_transp_short) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>% 
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0))

amces_transp_crackdown_tbl <- amces_transp_crackdown_medians %>% 
  ungroup() %>% 
  mutate(nice = glue("{fmt_decimal(y)}\n{build_ci(ymin, ymax)}")) %>% 
  mutate(feat_transp_short = str_replace(feat_transp_short, " - ", "−")) %>% 
  mutate(across(starts_with("p_"), ~fmt_decimal2(.))) %>% 
  arrange(desc(feat_govt)) %>% 
  select(contrast = feat_transp_short, diff_nice = nice, p_neq_0)

mms_acc_crackdown <- preds_all %>% 
  group_by(feat_acc_short, feat_govt, .draw) %>% 
  summarize(avg = mean(.epred))

mms_acc_crackdown_tbl <- mms_acc_crackdown %>% 
  group_by(feat_acc_short, feat_govt) %>% 
  median_qi() %>% 
  mutate(nice = glue("{fmt_decimal(avg)}\n{build_ci(.lower, .upper)}")) %>% 
  ungroup() %>% 
  arrange(desc(feat_govt), desc(feat_acc_short)) %>% 
  select(feat_acc_short, feat_govt, nice)

amces_acc_crackdown_medians <- mms_acc_crackdown %>% 
  group_by(feat_govt) %>% 
  compare_levels(variable = avg, by = feat_acc_short, comparison = "control") %>% 
  group_by(feat_govt, feat_acc_short) %>% 
  summarize(
    median_qi(avg),
    p_gt_0 = sum(avg > 0) / n()
  ) %>% 
  mutate(p_neq_0 = ifelse(y >= 0, p_gt_0, 1 - p_gt_0))

amces_acc_crackdown_tbl <- amces_acc_crackdown_medians %>% 
  ungroup() %>% 
  mutate(nice = glue("{fmt_decimal(y)}\n{build_ci(ymin, ymax)}")) %>% 
  mutate(feat_acc_short = str_replace(feat_acc_short, " - ", "−")) %>% 
  mutate(across(starts_with("p_"), ~fmt_decimal2(.))) %>% 
  arrange(desc(feat_govt)) %>% 
  select(contrast = feat_acc_short, diff_nice = nice, p_neq_0)

trans_acc_crackdown_tbl <- bind_rows(
  bind_cols(
    mms_transp_crackdown_tbl, 
    amces_transp_crackdown_tbl %>% 
      add_row(contrast = NA, .after = 1) %>% 
      add_row(contrast = NA, .after = 3) %>% 
      add_row(contrast = NA, .after = 5)
  ) %>% 
    mutate(feat_level = paste("Transparency:", feat_transp_short)) %>% 
    select(-feat_transp_short),
  bind_cols(
    mms_acc_crackdown_tbl, 
    amces_acc_crackdown_tbl %>% 
      add_row(contrast = NA, .after = 1) %>% 
      add_row(contrast = NA, .after = 3) %>% 
      add_row(contrast = NA, .after = 5)
  ) %>% 
    mutate(feat_level = paste("Accountability:", feat_acc_short)) %>% 
    select(-feat_acc_short)
) %>%
  mutate(contrast = ifelse(is.na(contrast), "*(Reference)*", contrast)) %>% 
  relocate(feat_level) %>%
  mutate(feat_govt = fct_relabel(feat_govt, ~paste("Relationship with government:", .x))) %>% 
  arrange(desc(feat_govt))

inner <- glue("
rowsep=1.5pt, row{<<header_rows>>}={abovesep=6pt, belowsep=0pt},
cells={font=\\footnotesize}, row{1}={valign=b}
",
  header_rows = paste0(get_row_group_index(trans_acc_crackdown_tbl$feat_govt) + 1, collapse = ","),
  .open = "<<", .close = ">>"
)

notes <- list(
  `*` = list(i = 0, j = c(2, 4), text = tt_md(paste0("N = ", details$combos, " estimated marginal means collapsed into feature-specific means. Values are on the percentage-point scale; single value is posterior median; 95% credible interval in brackets."))),
  `†` = list(i = 0, j = 5, text = tt_md("The probability of direction (*p*~direction~) is the probability that the posterior AMCE is strictly positive or negative—it is the proportion of the posterior AMCE that is the sign of the median."))
)

trans_acc_crackdown_tbl %>%
  select(-feat_govt) %>%
  mutate(across(c(nice, diff_nice), ~tt_linebreak(.))) %>%
  setNames(
    c("Feature", "Posterior EMM", "Contrast", "Posterior ∆", tt_md("*p*~direction~"))
  ) %>%
  tt(width = c(0.30, 0.18, 0.25, 0.18, 0.06), notes = notes) %>%
  group_tt(
    i = as.character(trans_acc_crackdown_tbl$feat_govt), 
    indent = 0
  ) %>%
  format_tt(replace = "—") %>%
  format_tt(j = 3, markdown = TRUE) %>%
  format_tt(j = c(2, 4, 5), escape = FALSE, replace = "—") %>%
  style_tt(j = c(1, 3), align = "l") %>%
  style_tt(j = c(2, 4, 5), align = "c") %>%
  style_tt(
    i = get_row_group_index(trans_acc_crackdown_tbl$feat_govt),
    # line = "b", line_width = 0.05, line_color = "#d3d8dc", align = "l", bold = TRUE,
    line = "b", line_width = 0.05, align = "l", bold = TRUE
  ) %>%
  style_tt(
    i = tail(get_row_group_index(trans_acc_crackdown_tbl$feat_govt), -1),
    # line = "t", line_width = 0.05, line_color = "#d3d8dc"
    line = "t", line_width = 0.05
  ) %>%
  style_tt(
    bootstrap_class = "table",
    tabularray_inner = inner
  )
```

\clearpage

::: {.content-visible when-meta="appendix-single-spacing"}
\singlespacing
:::

# Preregistration deviations

We made the following deviations from our preregistered protocol [@WillrothAtherton:2024]:

1. Type

   :   Analysis

   Reason

   :   New knowledge

   Timing

   :   After results known

   Original wording

   :   "We will examine the aggregate marginal posterior distributions of the attribute levels"

   Deviation description

   :   This statement was vague and seemed to imply that we would analyze the results of the model by looking only at the raw model coefficients. While is is possible to calculate exact feature contrasts by summing specific combinations of coefficients, we instead calculated estimated marginal means and their contrasts (or AMCEs) using the fitted model.

   Reader impact

   :   This deviation should improve readers' interpretation of the findings, since the reported results are no longer on a log-odds or logit scale, and instead are on a more interpretable percentage point scale—estimated marginal means show the percent of respondents who support an NGO given a specific features, while AMCEs show the percentage point change in support when moving from one feature to another. The risk of bias is minimal as the underlying results are identical whether reported as logit-scale coefficients or marginal means.

2. Type

   :   Hypotheses

   Reason

   :   Stylistic

   Timing

   :   After results known

   Original wording

   :   Q5a: "Donors will show increased willingness to donate to NGOs that are financially transparent"

   Deviation description

   :   We rephrased this as H~1a~: "If NGOs are financially transparent, then individual private donors will have a higher likelihood of supporting or donating to them."

   Reader impact

   :   This deviation has minimal impact on readers' interpretation of the findings—it is only rephrased to follow an "if… then…" formulation.

3. Type

   :   Hypotheses

   Reason

   :   New knowledge + stylistic

   Timing

   :   Direction restated after data collection, but before results were known; "if… then…" formulation added after results known

   Original wording

   :   Q5f: "Donors should be no more or less likely to donate to NGOs that are accountable and hold regular third party audits"

   Deviation description

   :   We rephrased this as H~1b~: "If NGOs are accountable, then individual private donors will have a higher likelihood of supporting or donating to them."

   Reader impact

   :   This deviation might have some impact on readers' interpretation of the findings. This deviation was the result of misunderstanding existing work on the effect of nonprofit accountability on donor behavior, and we hypothesized that there would be no effect, contrary to what is predicted by previous research. The risk of bias is low, however—we reversed our prediction after data collection but before we analyzed the data and before the results were known.

4. Type

   :   Hypotheses

   Reason

   :   Stylistic

   Timing

   :   After results known

   Original wording

   :   Q2a: "Donors will show increased willingness to donate to NGOs that are facing government crackdown or criticism"

   Deviation description

   :   We rephrased this as H~2~: "If NGOs face legal crackdowns abroad, then individual private donors will have a higher likelihood of supporting or donating to them."

   Reader impact

   :   This deviation has minimal impact on readers' interpretation of the findings—it is only rephrased to follow an "if… then…" formulation.

5. Type

   :   Hypotheses

   Reason

   :   Stylistic

   Timing

   :   After results known

   Original wording

   :   Q5b: "Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent"

   Deviation description

   :   We rephrased this as H~3~: "If NGOs face legal restrictions abroad *and* are financially transparent, then individual private donors will have a higher likelihood of supporting or donating to them."

   Reader impact

   :   This deviation has minimal impact on readers' interpretation of the findings—it is only rephrased to follow an "if… then…" formulation.

6. Type

   :   Hypotheses

   Reason

   :   New knowledge

   Timing

   :   Accountability prediction added after data collection, but before results were known; "if… then…" formulation added after results known

   Original wording

   :   Q5b: "Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent"

   Deviation description

   :   We explore the interaction between (1) government crackdown and financial transparency and (2) government crackdown and accountability in the paper, but we only specified the first interaction in the preregistration.

   Reader impact

   :   This deviation might have some impact on readers' interpretation of the findings. The omission of a prediction of the relationship between government crackdown and accountability was inadvertent and we had intended to specify it. The risk of bias is low, as we added the new crackdown+accountability hypothesis after data collection and before the results were known.

7. Type

   :   Hypotheses

   Reason

   :   Narrative

   Timing

   :   After data collection, before results were known

   Original wording

   :   Q1: Branding; Q3: Issue area; Q4: Funding sources

   Deviation description

   :   For the sake of narrative simplicity, we do not explicitly test these three predictions as hypotheses. In this paper, our primary interest is crackdown, transparency, and accountability, but we look at branding, issue area, and funding sources to help compare and give context to the magnitude of our main hypotheses.

   Reader impact

   :   This deviation might have some impact on readers' interpretation of the findings, as it might appear that we have selectively reported a handful of our predictions. To avoid this, and for the sake of full transparency, we include these results in Figure 1 in the paper and @tbl-all-results. The risk of bias is low—we decided on the narrative framing for this paper after collecting the data but before analyzing the results.

\newpage

# Condensed preregistration

::: {.callout-important icon=false}
### Consensed preregistration

This is an anonymized and condensed version of our full OSF preregistration protocol, which is available at <https://doi.org/10.17605/OSF.IO/HSBYD>.
:::

{{< include text/preregistration.md >}}

\newpage

# References